### 2021-10-03 article/article.Rmd
text edits, article and appendix

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: 
  - Gabriel Cepaluni[^gabriel]
  - Danilo Freire[^danilo]
  - Umberto Mignozzetti[^umberto]
date: \today
thanks: "We thank Cedric Antunes, Victor Araujo, Eric Arias, Taylor Boas, Natalia Bueno, Andrea Freitas, Saad Gulzar, James Mahon, Lucas Novaes, Nelson Ruiz, Catarina Roman, David Skarbek, and Denis Stukal for their valuable comments. We also thank the participants of seminars and conferences at CEPESP, CUNY, INSPER, IPSA, LACEA, LASA, MPSA, Notre Dame, NYU, PUC Chile, SPSA, and Uppsala. We gratefully acknowledge the financial support of the São Paulo Research Foundation (FAPESP), grant number 2017/07290-5. All materials required to replicate this study are available at <https://github.com/umbertomig/legislature-size-welfare>."
abstract: "How does legislature size impact public service provision? Despite the importance of institutional design for democratic governance, the effect of particular legislative features on citizen welfare remains little understood. In this paper, we use a formal model to show that increasing legislature size improves the provision of public goods. We argue that changes in bargaining costs depend on whether additional legislators share the Executive's party affiliation: More opposition members reduce the Executive's public goods provision, while more government-aligned members increase it. We test this theory by exploiting sharp discontinuities in city council size in Brazil. We show that an additional city councilor has a 91% chance of belonging to the mayoral coalition, and this significantly improves primary school enrollment and infant mortality measures. To explore possible mechanisms, we surveyed former councilors and analyzed 346,553 bills proposed between 2005 and 2008. This paper has implications for the design of representative institutions."
abstractspacing: double
keywords: Legislature Size; Legislative Bargaining; Local Public Goods Provision; Regression Discontinuity Design; Brazil
jelcodes: H41; H80; I31; I38
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
# Knitr options
knitr::opts_chunk$set(fig.pos = "H") # holds figure position
knitr::opts_chunk$set(echo = TRUE)
```

[^gabriel]: Associate Professor, Department of International Relations, São Paulo State University, <gabi.cepal@gmail.com>.

[^danilo]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <dfreire@lincoln.ac.uk>, <https://danilofreire.github.io>.

[^umberto]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>. Corresponding author.

\newpage

# Introduction
\label{sec:intro}

Legislative institutions are crucial for citizen welfare and service provision in democratic countries. Historically, the division of power and the birth of a representative government placed legislatures at the center of the political stage. In modern democracies, most countries have assemblies at all levels of government. These institutions regulate the quantity and quality of legislation, check and veto other political powers, and influence policymaking. Moreover, legislatures also determine taxation, government spending and oversee public programs \citep{auriol2012optimal, north1989constitutions, poulsen2018political, stasavage2011states, Weingastetal1981, weingast1988industrial}.

Regardless of eventual differences in electoral rules, social composition, and internal procedures, all legislatures share one feature: they are collective bodies. As such, the number of representatives in parliament has important consequences for whole polity. Indeed, legislature size is known to have a significant effect on collective action strategies, the bargaining power between state branches, government spending, and representation \citep{allenrepresentation, crain1979cost, freire2021legislature, PrimoSnyder2008, rogers2002free, PrimoSnyder2008, Weingastetal1981}. However, how legislature size impacts social welfare remain less understood. On the one hand, enlarged legislatures may increase government spending and lawmaking, which can lead to more public goods \citep{Weingastetal1981}. On the other hand, adding a legislator may increase the number of veto players and exacerbate collective action problems, thus hindering service provision \citep{gottlieb2019countervailing, tsebelis2002veto}. Therefore, what is the effect of larger legislatures on public service provision?

In this paper, we propose a local legislature size and welfare theory that incorporates two crucial features: Executive leadership and party politics. First, in mayor-council municipalities, mayors tend to be central in local public service decisions. Mayors have the prerogative to boost public investments, like presidents in federal governments. Second, partisanship is essential at the local level, even in settings where the ideological attachment is weak, such as in most young democracies. Party attachment structures political views \citep{samuels2014power}, and local-level elites usually belong to different factions and parties. They fight for power in the municipality by controlling various parties and forming distinct pre-electoral coalitions.\footnote{Elite political competition in the local level has also been linked to violence and conflict in Brazil and elsewhere \citep{hoelscher2015politics, mignozzetti2018predatory}.} In many cities, family ties may be substitutes for the ideological component, but influential local leaders usually sort themselves into different political groups \citep{ferraz2020political, nunes1977coronelismo}.

This paper incorporates partisanship and mayoral first-move advantage to a \citet{baron1989bargaining} bargaining model. Mayors make the initial policy proposal that consists of an investment in public goods provision, a decision to extract rents for personal gains, and a vector of side-transfers for councilors. If they fail to secure majoritarian support from councilors, a reversal stage occurs, where the decision leaves the mayor's office and councilors propose the public investments. In this mayor-council bargaining process, costs come from the side-transfers required to secure a majority among councilors. We show that in a model without partisan concerns, bargaining costs always increase with larger city councils, which in turn lower public goods provision and rents. However, partisanship decreases bargaining costs when the chances of electing councilors aligned with the mayor are sufficiently high. Lower bargaining costs release funds for mayors to invest in public goods provision. In this case, a larger city council may have more partisan councilors aligned with the mayor, reducing the mayor's bargaining costs and increasing the equilibrium public goods provision.

We test these theoretical predictions by exploiting an exogenous variation in Brazilian city council sizes. In 2004, the judiciary reinterpreted the Brazilian Constitution and unexpectedly changed the number of councilors in several municipalities. Until then, towns could freely determine their legislature size, but in March 2004 the Supreme Electoral Court created population thresholds and added one councilor to each set of 47,619 inhabitants. Around these cutoffs, the decision represented an exogenous change in city council size, allowing us to investigate the effects of increasing the number of local legislators in Brazil.

We find that larger city councils changed the composition of mayoral coalitions and the number of appointed bureaucrats in the municipality. The extra city councilor had a 91 percent chance of belonging to the mayor's pre-electoral coalition. This enhanced the chances of electing a councilor connected with the mayor, which then decreased the bargaining costs for the mayor and boosted investments in public services. Moreover, the number of appointed bureaucrats increased by 104 with the additional councilor. In Brazil, appointing bureaucrats leads to higher quality workers, better service provision, and more accountability \citep{toral2019benefits}.

Higher mayor support and more appointed bureaucrats correlate with city councilors' service provision expectations and legislation approved in Brazilian municipalities. We surveyed 174 former councilors who served during the 2005-2008 term, and they corroborate that mayors use bureaucratic appointments and political favors to secure support. As the formal model shows, bargaining with councilors requires offering transfers to the representatives. By investigating what councilors prefer to provide to voters, we find that councilors mainly demand job appointments, as these increase their reelection chances and improve municipal welfare. These findings align with a novel dataset we built with 346,553 laws passed in 63 municipalities within ten thousand inhabitants away from the cutoffs. The legislation in municipalities on the left and on the right-hand side of the cutoffs mentioned local public goods. However, municipalities with larger city councils had around 15 percent more public goods provision proposals, what is in line with our expectations.

Finally, we show that the increase in legislature size significantly impacts municipal education and health care. Increasing the city council size by one legislator lowers infant mortality by 2.01 children per one thousand infants born and reduces postneonatal mortality by 0.90 children per one thousand infants that survived their first 28 days. A larger council size also increased enrollment by 2.58 children in elementary education classrooms without significantly diminish elementary school quality. These services are crucial for welfare and result from an eased bargaining between the mayor and the city council.

This paper contributes to three strands of the literature. First, this paper dialogues with a large literature on multilateral bargaining. Repeated models of multilateral bargaining started with \citet{baron1989bargaining}, who extended the bilateral bargaining game proposed by \citet{rubinstein1982perfect} to more than two players. Several papers followed \citet{baron1989bargaining}, manipulating endogenous recognition power, the size of the required majority, and partisanship in the models \citep{agranov2020persistence, baranski2020determinants, calvert2005legislative, choate2019partisan}.\footnote{See \citet{eraslan2019legislative} for a comprehensive review of multilateral bargaining and \citet{ray2015coalition} for a review on coalition formation.} In this paper, we extend the \citet{baron1989bargaining} model under closed rules and add the mayor's first move and political partisanship. Partisanship here is different from \citet{choate2019partisan} since we consider partisanship as a zero-sum effect that is positive for politicians aligned with the mayor and negative otherwise. This paper investigates the often overlooked effect of legislature size on the bargaining equilibrium strategies.

Second, this paper contributes to the literature on the political and economic effects of larger legislatures. Following the contribution by \citet{Weingastetal1981}, the "law of $1/n$" posits that larger legislatures increase the legislative pressure for providing targeted public goods to local supporters at the expense of the entire polity \citep{freire2021legislature}. The corollary of the "law of $1/n$" is that increasing legislature size promotes excessive public spending.\footnote{\citet{PrimoSnyder2008} show that legislature size may result in positive or negative expenditures, depending on factors such as the degree that local public goods can exclude non-locals or deadweight costs associated with taxation.} In the US metaphor, this would explain significant infrastructure spending in smaller districts. However, the "law of $1/n$" literature rarely mentions the relationship between public goods provision and welfare. This paper is one of the first to study how the number of local legislators affects welfare. Our model provides a theoretical expectation based on the chances of electing a politician aligned with the mayor. The empirical estimates confirm our theoretical predictions, with the additional fact that legislature size has a null effect on expenditures in Brazil.

Finally, our article also contributes to a growing literature on municipal partisan politics. Most papers in this area estimate the effects of the mayor's party on political outcomes. For instance, \citet{gerber2011mayors} and \citet{de2016mayoral} find differences in governance when comparing Democratic and Republican mayors. Specifically about Brazil, \citet{gouvea2021partisanship} shows that social policy expenditure increase with left-wing mayors. \citet{frey2020larger}, using discontinuities in city council size after the 2009 constitutional amendment that reverted the thresholds we study here, shows that changes in council size lower the support for candidates backed by the municipal mayors. This paper reinforces the view that partisanship matters locally, even in countries with weak ideological linkage.

# A Theory of Legislature Size, Legislative Bargaining, and Welfare
\label{sec:theory}

Several cities in the world are governed by the mayor-council system. In mayor-council cities, the power is divided between a mayor that controls the Executive branch and a city council that supports or sanctions the policies proposed by the mayor. In this setting, consider an strategic interaction between a mayor ($M$) and $N \geq 3$ city councilors indexed in $i \in \{1, 2, \cdots, N \}$. The mayor has the prerogative of proposing a vector of policies that is voted by the city councilors. The provision of public goods is then a bargain between the mayor and the city council. Policy proposals are a combination of public goods provision $g$, rents for the mayor $r$, and a vector of transfer for city councilors \textbf{x}.

The city council votes the mayor's proposal, and if councilors accept it, the proposal is implemented. Otherwise, a reversal policy takes place. Reversal policy in this context means that the mayor's proposal failed to secure the majority of council votes. The decision process then leaves the mayor's office and it is transferred to the council. The types of reversal policies are crucial for our argument, as it changes the relative strength of the council when they bargain with the mayor. We investigate two types of reversal policies: a baseline reversal with no parties and a hybrid mechanism with partisan and non-partisan concerns. The differences between these two mechanisms are that the non-partisan councilors only care about rents, while partisan councilors care about both rents and policy.

The mayor must convince at least half of the city councilors to support her policy proposals. Similar to \citet{buenodemesquita2005logic}, we assume that the selectorate, that is, the individuals who have a say in the mayor's policy proposals, are the $N$ city councilors. The winning coalition, the minimum number of members that must be convinced to implement the policy, is equal to half of the selectorate ($N/2$). City councilors are motivated by policy concerns $p$ and transfers received from the mayor or the councilor recognized as the proposer \textbf{x}. We leave the definition of these transfers open, as this allows us to analyze how mayors combine transfers with other incentives to pressure city councilors. Exemples of transfers are granting electoral favors, favoring clientelistic brokers, spending in areas that the councilor has political or personal interests, providing portfolio (municipal offices) and public jobs within the municipal bureaucracy, or even giving campaign contributions and bribes. The use of transfers to influence the councilors' choices generate a bargaining cost of $C_G$. Bargaining costs vary according to the reversal mechanism and the size of the legislature.

Finally, after learning the bargaining costs, the mayor provides a level of public goods provision $g$ and pockets rents $r$. Public goods provision helps the mayor to get reelected. Rents are for direct mayor's consumption, and they do not contribute to the mayor's electoral success.\footnote{This model, similar to \citet{cowen1994rent} and \citet{buenodemesquita2005logic}, combines public goods provision and rent extraction. For a dynamic bargaining model that incorporates transfers to districts and public goods provision, see \citet{battaglini2012legislative}.} This makes the mayor's expected utility a sum of the gains from rents and the benefits from reelection. Both the utilities from rents $u(.)$ and the probability of reelection $\pi(.)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate. The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains that the mayor receives from holding the public office. We assume that the benefits are high enough to rule out an equilibrium with only rents and transfers, which is equivalent to saying that mayors strongly prefer to be reelected. If the mayor's policy proposal is approved, then the expected utility for the mayor is:

\[
\mathbb{E}U(r, g) \ = \ u(r) + B_M\pi(g)
\]

The policy choices of the mayor are subject to municipal budget constraints. The municipality has $R > 0$ resources and cannot run debts, meaning that the budget must be balanced. We also assume that there are enough resources for the mayor to govern.

Let $C_G$ be the bargaining costs. The budget balance constraint requires that the offers the mayor makes to the city councilors must satisfy the following inequality:

\[
r + g + C_G \ \leq \ R
\]

The expected utility for the city councilors depends on the reversal policy. In the baseline non-partisan reversal, when the council rejects the mayor's proposal, a reversal stage starts with resources diminishing by a factor of $\delta \in (0,1)$ and the selection of one councilor randomly. The selected councilor becomes the new proposer. If her proposal is accepted, it is implemented. If councilors reject the offer again, the budget shrinks by $\delta$, and another city councilor is recognized to propose. The process repeats until a proposal is finally accepted.

We now consider a situation that combines both partisan and non-partisan legislature concerns. Each councilor has a party affiliation. Party affiliations are mutually exclusive, and if a city councilor is aligned with the mayor's party, then she belongs to the government coalition $G \subset \{1, 2, \cdots, N \}$. Otherwise, the city councilor belongs to the opposition $O = \{1, 2, \cdots, N \} \setminus G$. The mayor's policy generates a political value of $p > 0$ for the councilors, and whether it adds or subtracts from the councilor's utility depends on the party affiliation of the councilor. If councilors reject the mayor's proposal, no public goods are provided ($p=0$). Moreover, in equilibrium, we assume that city councilors would prefer to receive transfers instead of providing public goods.\footnote{In the Supplementary Materials, we consider a version of the model where public goods provision depends on the party affiliation of the proposer so that it affects choices in any potential history of the game. The equilibrium public goods provision level is even higher than in the model presented here, as city councilors prefer to approve the mayor's proposal than run the risk of an opposition member becoming the proposer.}

The timeline of the strategic interaction is as follows:

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, x)_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts the proposal, the policy is implemented and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) The budget is discounted by a factor $\delta \in (0,1)$, and one councilor is randomly selected to make an offer. If the offer is accepted, it is implemented. If rejected, the reversal policy stage restart.\footnote{In section A.1 of the Supplementary Materials, we solve three reversal mechanisms. The two we present here, plus a mechanism with only political concerns.}
\end{enumerate}

The solution concept we use in this model is the subgame perfect Nash equilibrium. This game requires that the strategies follow a Nash equilibrium in each subgame. We find the solution using backward induction. In the model with infinitely repeated proposals, any strategy can be shown to be a Nash equilibrium. Thus, a natural assumption is to extend the equilibrium concept to require the equilibrium to be stationary. A stationary subgame perfect equilibrium requires that, at each given point of the game, if a politician accepts an offer at time $k+1$, she should take the same offer at time $k$ \citep{rubinstein1982perfect}. Stationarity gives us a method to find a proposal that would be accepted at any stage of the game. Hence, we can characterize a sequence of offers for each point $k$ in time. For the mayor, the optimal offer is at $k=0$, representing a no delay in policy implementation. Moreover, we assume that the players play no dominated strategies, even when there is insufficient support for the policy.

For the solution to be of general interest, we solve the bargaining costs for the expected number of opposition and government-aligned city councilors. This is because the solution for each fixed number of opposition and government councilors does not help us understand the overall relationship between council size and bargaining costs.

## The Mayor's Decision Stage

Solving the game by backward induction requires us to start with the councilors' strategies. However, to conserve space, assume that we solved the game for the councilors and found the equilibrium bargaining costs $C_G(N)$ when there are $N$ city councilors. We now derive the optimal rents ($r$) and public goods provision ($g$) proposed by the mayor. The mayor benefits from public goods provision, as it increases the chances of her reelection. However, she prefers to invest as minimum as necessary in public goods and extract the remaining resources as political rents. The mayor's objective is to maximize her expected utility, subject to the municipal budget constraint.

\[
\begin{aligned}
\max_{r, g} \quad & u(r) + B_M\pi(g)\\
\textrm{s.t.} \quad & r + g + C_G(N) \leq R
\end{aligned}
\]

In equilibrium, the budget binds, and the first-order condition for an optimal public goods provision makes the marginal benefits to increase rents. In this case, it means that the marginal costs of providing the public goods is equal to the marginal benefits of reelection:

\[
u'(R - g - C_G(N)) = B_M\pi'(g)
\]

The optimal level of public goods provision equates the marginal gains accrued from rents with the marginal benefits from reelection.

\begin{proposition}\label{prop1}
The provision of public goods increases with legislature size when the bargaining costs decrease with legislature size.
\end{proposition}

All proofs are in section A.1 of the Supplementary Materials. The proof's intuition involves finding what happens to the mayor's marginal utility when we increase both $g$ and $N$. As $N$ increases discretely on a complete lattice, we use monotone comparative statics to derive the conditions for increasing differences \citep{milgrom1994monotone, ashworth2006monotone}. Increasing differences mean that higher $N$ leads to higher equilibrium $g$.

Proposition \ref{prop1} provides the first empirical hypothesis to test: \textit{if the bargaining costs decrease when the legislature size increases, then the provision of the public goods increases when the legislature size increase}. Now we study each reversal mechanism to determine the condition under which the bargaining costs decrease when legislature size increases.

## Baseline: Non-Partisan Legislative Bargaining

In the non-partisan reversal, we assume that when the council rejects the mayor's office, a reversal stage starts with the random selection of one councilor. This formulation is similar to the closed rules proposal in \citet{baron1989bargaining}. To find a stationary subgame perfect Nash equilibrium, suppose that there were $k-1$ rejections, and the game is at the $k$-th stage. A councilor accepts the proposer's offer if, and only if, to accept the offer at $k$ is better than or equal to wait until the next stage $k+1$. If the offer is $x_i$, then:

\[
x_i \geq \dfrac{1}{N}\left[ \delta^{k+1}R - \dfrac{N}{2}x_i \right] + \left[ 1 - \dfrac{1}{N} \right]\left[\dfrac{1}{2} x_i \right]
\]

We place the offer on the left-hand side. There are two components on the right-hand side. The first is the amount that the councilor $i$ gets when she is the proposer. It is equal to the chance that she is recognized as the proposer times the budget in the next round minus the offers she makes to convince half of the councilors. The second part represents the gains if the councilor rejects the current received offer but still gets an offer in the following round. It is equal to the chance that she is not recognized as the proposer times the chance that she receives a transfer times the transfer amount. Note that the offer the councilor makes as the proposer is the same as the offer she wants to receive. This is because the city councilors are exchangeable. The solution is symmetric for all councilors receiving an offer (this means that, without loss of generality, we could have dropped the $i$ in the solution). After some algebra, the offer $x_i$ must be greater than or equal to:

\[
x_i \geq \dfrac{2 \delta^{k+1}R}{2N + 1} \equiv \underline{x}(k, N)
\]

The proposer always offers the minimum required to get the proposal approved. In this case, the offer at any given stage $k$ is going to be equal to $\underline{x}(k, N)$.

Proceeding backwards, at the mayor's proposal stage $k=0$ the mayor is going to offer $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the councilors. In this context, the bargaining costs for the mayor are:

\[
C_G(N) \ = \ \dfrac{N}{2}\left[\dfrac{2 \delta R }{2N + 1}\right] \ = \ \dfrac{\delta R N}{2N + 1}
\]

\begin{proposition}
In the baseline non-partisan reversal, bargaining costs always increase when the size of the legislature increases.
\end{proposition}

The intuition for the results is that every additional councilor demands more resources from the mayor. Thus, more councilors represent higher bargaining costs. And if the bargaining costs increase with legislature size, then any increase in public goods provision associated with legislature size has come from a different mechanism. Below we show that partisanship is the key to understand how larger legislature sizes can increase public goods provision.

## Partisan and Non-Partisan Bargaining

In a reversal mechanism that incorporates partisan and non-partisan motivations, a city councilor aligned with the government favors the mayor's offer if:

\[
x_i \geq \dfrac{2\delta R}{2N+1} - p
\]

An opposition politician, on the other hand, favors the mayor's offer if, and only if:

\[
x_i \geq \dfrac{2\delta R}{2N+1} + p
\]

We decompose these costs into two components. First, the costs in terms of rents. Second, the costs (or benefits) from political alignment. In this context, the bargaining costs depend on the mayor's offers for councilors in her coalition versus the mayor's proposals for councilors in the opposition. As the ex-ante chance of an opposition member being elected is equal to $1-\gamma$, after taking the weighted averages these costs become:

\[
C_G(N) \ = \ \dfrac{N}{2}\left( \dfrac{2\delta R}{2N+1} - p \right) + \left(\dfrac{N}{2} - \gamma N \right)2p
\]

\begin{proposition}
In the reversal mechanism that incorporates partisanship, if $\gamma \geq \dfrac{1}{p}\left[ \dfrac{1}{(2N+1)(2N+3)} \right] \equiv \underline{\gamma}$, then the bargaining costs decrease when the legislature size increases.
\end{proposition}

The intuition for the proof is that as partisanship influences the costs and benefits of legislators, more government-aligned councilors make it cheaper to govern. When the partisan benefit is higher than the additional cost that a legislator generates, adding a councilor will benefit mayors if the councilor has a higher chance of sharing the same party as the mayor. Therefore, a higher probability of electing a legislator aligned with the mayor reverts into lower bargaining costs.

Also, the threshold $\underline{\gamma}$ that captures the probability of electing a mayor-aligned policitian decreases along with the political value of policy $p$ and the size of the legislature $N$. This means that intensive partisanship and larger legislatures make it easier to satisfy the electoral threshold.

## Main Hypotheses

Our model provides two empirically testable hypotheses. Consider a municipality with a city council comprised of $N$ legislators. Then the city increases the size of the council to $N+1$. Thus, in a partisan setting:

\begin{enumerate}[labelindent=*, 
style=multiline, leftmargin=*, 
label=H\arabic*.]
  \item Bargaining costs decrease with legislature size when the chances of electing a government-aligned legislator are sufficiently high.
  \item Public goods provision increases if bargaining costs decrease with legislature size.
\end{enumerate}

## Competing Hypotheses

In this paper, we show that the provision of education and health care in Brazil improves with larger legislatures. This is because the chances of electing a city councilor aligned with the mayor are 91 percent, thanks to the rule of distributing seats adopted in the country.

However, other pathways may also explain the nexus between city council size and welfare. First, improvements in representation could change the provision of health care and education. Increasing legislature size without changing the population makes councilors represent fewer people. For example, in a town with 50 thousand voters and ten councilors, each councilor represents five thousand voters. If the same city has eleven councilors, then each councilor represents 4.5 thousand voters. This facilitates the representation of minorities, as they have to gather fewer votes to elect a city councilor. A minority representative can advocate for policies that would improve the welfare of marginalized groups \citep{ChattopadhyayDuflu2004, chin2011redistributive, pande2003can}. Diversity could also improve governance in municipalities \citep{rugh2011provision}. For instance, more women in public offices may shift service provision toward welfare expenditure \citep{brollo2016happens, duflo2012women, funk2019representative, hernandez2018women}. Therefore, an alternative hypothesis to explain the increased service provision is that larger legislatures improve female and minority representation in Brazil.\footnote{For a description of the barriers faced by non-white politicians in Brazil, see \citet{janusz2021race}.}

Second, more seats could increase the competitiveness in the election, making politicians provide more services \citep{arvate2013electoral}. In a political system with around thirty parties and a sample average of almost ten parties in the municipalities we study, candidates per seat can increase exponentially. According to the Brazilian electoral law, each party can sponsor 1.5 times the number of seats as candidates. If a party form a pre-electoral coalition, this number increases to two times the number of seats \citep{coneglian2008lei}. An election in a small town, with ten councilors, may end up with 600 candidates: (30 parties) x (two times the number of seats) x (ten seats). Therefore, an alternative hypothesis to explain the increased service provision is that larger council size increase the competitiveness of the election.\footnote{Despite the common assumption that competitiveness improve welfare, \citet{boulding2014political} and \citet{gottlieb2019countervailing} can increase budgetary pressure, lower social policy, lead to bargaining inefficiencies.}

Finally, the mayor and the legislature carry out the provision of services. Similar to what happens in firms, an extra employee can increase the firm productivity frontier by improving cooperation and synergistic outputs \citep{alchian1972production, holmstrom1982moral}. A similar argument would posit that the increase in council size increases the public service production capability in the municipality. As the production of the legislature manifests itself into legislation, an alternative hypothesis to explain the increased service provision is that more laws proposed by the city councilors are being approved.

\begin{enumerate}[labelindent=*, 
style=multiline, leftmargin=*, 
label=AH\arabic*.]
  \item Female representation increases with legislature size, improving public goods provision.
  \item Non-white representation increase with legislature size, improving public goods provision.
  \item Election competitiveness increases with legislature size, improving public goods provision.
  \item Legislative production increases with legislature size, improving public goods provision.
\end{enumerate}

We show below that hypotheses H1 and H2 hold, while the data falsify the alternative hypotheses AH1 to AH4.

# Background on the Brazilian Case
\label{sec:background}

In 2004, Brazil comprised 5,560 municipalities. According to the Brazilian Constitution, each city must provide health care, primary education, transportation, and infrastructure to its citizens. Municipalities have the authority to enact local laws and collect taxes on housing and services. However, their room for fiscal manipulation is limited. On average, municipal taxes account for only 5\% of the local budgets, while mandatory transfers from state and federal governments correspond to the remaining 95\%. This reduces the ability of municipal legislators to effectively increase local spending, which makes improvements in local services dependent only on the composition of their local investments.

With regards to their political organization, municipalities have one mayor and one city council. Citizens choose their representatives by direct vote in the same election. Mayors oversee the provision of public services and propose laws and amendments to the tax code. Proposals submitted by the mayor are subject to the approval of the city council. Councilors, in turn, have two primary duties. First, they should discuss and vote on legislation put forward by the mayors or by their peers. Second, they oversee the public budget and sanction mayors who do not comply with fiscal legislation. Councilors and mayors also provide an array of public and personal services for their constituents, such as helping voters access public hospitals, facilitating school enrollment, and even paying their voters' utility bills \citep{lopez2017legisladores, nichter2011electoral}.

Until 2004, there was no specific regulation on the size of city councils. The 1988 Brazilian Constitution set only broad guidelines about the number of council seats in each municipality, which allowed cities to determine how many legislators they should have in their local chambers. This led to severe imbalances in the proportion of representatives per citizen across the country. For instance, the Nova Russas municipality had 30,009 inhabitants and 21 councilors, accounting for one legislator per 1,429 citizens. In contrast, the city of Sorocaba had only 15 councilors for 528,735 inhabitants, that is, one legislator per 35,249 citizens. In 2003, this malapportionment problem gained national visibility when Mira Estrela, a small municipality with only 2,651 residents, reduced its city council from 11 to 9 seats because of fiscal considerations. The change motivated a legal dispute that reached the Brazilian Supreme Court. The Court favored the city's decision and ruled that the size of every local council should correspond to precise population thresholds.

Following the ruling of the Supreme Court, in March 2004 the Supreme Electoral Court established a series of population cutoffs for all municipal legislatures. Following the Brazilian Constitution, the Court specified that cities with up to one million residents should have between 9 and 21 legislators. The number of seats was allocated as follows. Each town starts with nine councilors, then adds one legislator for every 47,619 inhabitants until their councils reach 21 members. This threshold implies that cities with a population between 571,428 and one million should all have the same number of legislators, that is, 21 of them. Conversely, cities between one and four million inhabitants should have at least 33 councilors, and this number increases up to 41 members following the previous formula of one additional legislator per 47,619 inhabitants.

This decision makes Brazil the ideal testing ground for our theory. Although an endogenous decision motivated the Electoral Court ruling, the population cutoffs created sharp discontinuities in the distribution of local council seats. Cities close to the cutoffs had very similar characteristics, but those just above the threshold gained a new legislator. Moreover, cities could not self-select into any group, as their population estimates were calculated using the 2003 projections by the Brazilian Census Bureau (IBGE). This change in council size also happened only seven months before the election, making it impossible for candidates to adjust their service provision strategies for the 2004 election. Therefore, the ruling allows us to study the effect of larger legislatures on public service provision during the 2005-2008 term while holding all else constant.

# Variables and Data Sources
\label{sec:sources}

We use four groups of city-level variables in our models. The first group includes pre-treatment control variables that were not affected by the 2004 council size resolution. We add them to the estimates to improve the balance between treated and non-treated units. The covariates are: 1) the number of council seats in the previous term; 2) city population; 3) city GDP; 4) the proportion of low-income families in each municipality. The data come from the Supreme Electoral Court and the 2000 Brazilian census.

Second, we look into municipal data, city councilors' characteristics, approved legislation, and an online survey with 174 former city councilors that served during 2005--2008 term to investigate potencial political mechanisms.\footnote{We also analyzed 108 structured interviews conducted by \citet{lopez2017legisladores} at the Brazilian Economics Planning Institute (IPEA) in 2009. The authors asked councilors about their daily work. The interviews reinforced that the mayor's agenda dominates councilors' work and it is partisan without strictly following party lines. Municipal politics is divided into families or poles of power that belong to different party affiliations. Councilors from other parties gravitate around these groups, hence the opposition and government-aligned logic.} The municipal data contains: 1) the number of councilor's belonging to the mayor's pre-electoral coalition; 2) the number of appointed bureaucrats; 2) the councilor's gender and race; 3) the competitiveness of each city council seat; 4) the proportion of laws proposed by the city council that was approved. All variables come from the Supreme Electoral Court, the Senate legislative support Interlegis, and the Brazilian Bureau of Statistics.

Third, we also collected 346,553 bills from cities within ten thousand inhabitants from the 2004 population thresholds. We selected 63 out of 202 municipalities whose voting records were available online. We evaluate proposal types, discriminating between local public goods, oversight, and other legislative activity. Examples of the local public goods are investing in a public health clinic, requesting school staff to accommodate an extra child, renovating sewage, and fixing potholes on a given street.\footnote{Our theory does not differentiate between clientelism and broad service provision, as it is hard to disentangle differences in the local level legislation. A voter may request a particular service, such as fixing a pothole. Attending this request can be interpreted as a clientelistic practice. Still, the provision of the service generates non-excludable benefits to the adjacent neighborhood.} Oversight bills are accountability and transparency mechanisms meant to keep public services provision in check. Other legislation comprises bills that do not fit in the previous categories. Examples are changes in street names, motions to honor citizens or groups, and legislative and internal city council procedures. We provide further information about the legislation dataset and the city council survey in sections A.13 and A.14 of the Supplementary Materials.

The last set of variables quantify social welfare. Our focus is on the impact of legislature size on education and healthcare, which are the two most important public services the municipal government should provide to citizens. We measure access to education with the average enrollment in primary (K--4) municipal public schools. We assess education quality with the Education Development Index (\ 'Indice de Desenvolvimento da Educa\c{c}\~{a}o B\'asica--IDEB), which the Ministry of Education has issued biannually since 2007. The index is a weighted average of student scores in Portuguese and Mathematics (0--10), multiplied by the harmonic mean of public school promotion rates in a given year (0--100). To test the effect of the 2004 ruling on healthcare, we collect data on infant mortality per thousand children born and on postneonatal infant mortality, which is the death of infants who survived more than 28 days, but died before reaching one year of age. The source of all health care variables is the Brazilian Ministry of Health Data Center (DataSUS).

Together, this evidence provides a clear picture of local-level politics patterns in Brazil, from the political mechanism to the municipal welfare consequences of the larger city council size ruling. Summary statistics and variable descriptions follow in section A.2 of the Supplementary Materials.

# Empirical Strategy
\label{sec:empstrat}

We employ a regression discontinuity design (RDD) using the population thresholds established by the 2004 Supreme Electoral Court resolution as treatment indicators. Figure \ref{fig:runningvarplot} displays the distribution of municipalities within each city council size. These population thresholds remained in effect until 2009 when Congress amended the Brazilian Constitution. This study covers 2005 to 2008, which constitutes the full mayoral and city-council term following the 2004 Brazilian elections.

Our identification strategy relies on three assumptions. First, no municipality should be able to self-select into each side of the discontinuity. Second, as we estimate a sharp RDD, the council size should increase precisely as the law mandates. Third, this design assumes that the pre-treatment variables, collected before the 2003 Supreme Federal Court decision, were not affected by the 2004 ruling on council size. Moreover, for all the actual models estimated here, we fitted placebo cutoffs to check whether the estimation technique leads to consistent results. The placebo cutoffs are the dashed lines in Figure \ref{fig:runningvarplot}.

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=6in,angle=0]{../figures/distrSeats.pdf}}
      \caption{Distribution of Municipalities by Population and City Council Size}
      \bigskip
      \label{fig:runningvarplot}
    \end{minipage}
  \end{center}
\end{figure}

For the first assumption, in section A.3 of the online supplemental materials, we assess distribution imbalances running the \citet{McCrary2008} and \citet{cattaneo2019simple} tests. The tests show no distribution imbalances.

For the second assumption, running RDD in multiple thresholds may lead to inconsistent estimates when there are differences in the distribution of cases around the cutoffs \citep{cattaneo2016interpreting, eggers2018regression, bertanha2020regression}. In the dataset, there are twelve discontinuities. At each threshold, there is a sharp increase of one councilor, from 9 to 21. When running an RDD on these discontinuities, a consistent estimator should fit the exact change of one councilor. However, if we pool all the discontinuities together with no correction, we find an increase in council size of 1.63 councilors (see the first estimate in Table \ref{validitycheck}). This imbalance occurs because when we combine all the discontinuities, the model implies that municipalities below the 47,619 cutoffs (9 to 10 councilors) are comparable to changes in towns right above the 571,428 cutoffs (20 to 21 council members). That is incorrect because the comparison is inconsistent in council size, and these municipalities diverge in many other aspects.

To avoid this problem, in section A.5 of the Supplemental Materials, we show by simulation that adding controls, especially the variables responsible for the multiple threshold assignments (in our case, the population in 2003), improves the consistency and efficiency of the estimates. Relying on this fact, we add five controls to the estimates: population in 2003; GDP per capita; the number of seats in 2000; year; and a dummy for the northeast municipalities. The reason for the first variable is the multiple threshold assignments. We add GDP per capita because wealthier cities tend to have better public services. The number of seats in 2000 intends to control that some municipalities could experience a change provoked by the previous council size, confounding our estimates. The year and the Northeast dummies serve to improve efficiency. Adding controls improves efficiency in RDD and also enhances the consistency in the multiple thresholds RDD \citep{calonico2019regression}.

For the last assumption, there should be no changes in pre-treatment covariates. These pre-treatment covariates are variables measured before the primary outcomes. As the threshold rule selected by the Brazilian Electoral Court was unpredictable, it should not detect any variation before the thresholds were in place.

Table \ref{validitycheck} displays the results of the validity checks. We present the first-stage regressions at the top of the Table, with and without covariates, for both real and placebo cutoffs. At the bottom, we estimate the pre-treatment covariate balance. Furthermore, in the Supplementary Materials, we run placebo checks for all the models using placebo cutoffs within the real ones. The results are available in sections A.6 to A.12.

\begin{table}[htb!] \centering 
  \caption{Research Design Validity Check} 
  \label{validitycheck} 
    \scalebox{0.9}{
  \begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{4}{c}{\textbf{Panel A: Validity Check --- Treatment Effect on City Council Size}} \\ \cline{2-5}
   & \multicolumn{1}{c}{\small{(1)}} & 
 \multicolumn{1}{c}{\small{(2)}} & 
 \multicolumn{1}{c}{\small{(3)}} & 
 \multicolumn{1}{c}{\small{(4)}} \\

 & \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} & 
 \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} \\ 
  & \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} & 
  \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} \\
  & \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
\\[-1.8ex] 
 LATE & 1.63$^{***}$ & -0.31$^{**}$ & 1.00$^{***}$ & -0.00 \\ 
  & (0.51) & (0.14) & (0.0004) & (0.0000) \\ 
\hdashline \\[-1.8ex]   
 N Left & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} \\ 
Eff N Left & \multicolumn{1}{c}{199} & \multicolumn{1}{c}{638} & \multicolumn{1}{c}{49} & \multicolumn{1}{c}{886} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{385} & \multicolumn{1}{c}{51} & \multicolumn{1}{c}{477} \\ 
BW Loc Poly & 8.717 & 6.700 & 3.008 & 8.629 \\ 
BW Bias & 13.620 & 12.114 & 5.008 & 13.466 \\ 
\midrule 
 & \multicolumn{4}{c}{\textbf{Panel B: Validity Check --- Pre-Treatment Variables}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(5)}} & 
 \multicolumn{1}{c}{\small{(6)}} & 
 \multicolumn{1}{c}{\small{(7)}} & 
 \multicolumn{1}{c}{\small{(8)}} \\
  & \multicolumn{1}{c}{\small{Number of}} & 
 \multicolumn{1}{c}{\small{Population}} & 
 \multicolumn{1}{c}{\small{GDP}} & 
 \multicolumn{1}{c}{\small{\% of Poverty}} \\ 
  & \multicolumn{1}{c}{\small{Seats 2000}} & 
  \multicolumn{1}{c}{\small{2000 Census}} & 
  \multicolumn{1}{c}{\small{2000 Census}} &
   \multicolumn{1}{c}{\small{2000 Census}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
 LATE & 0.17 & -2057.60 & 0.06 & -9.72 \\ 
  & (0.67) & (1926.33) & (0.18) & (6.27) \\ 
  \hdashline \\[-1.8ex]
 N Left & \multicolumn{1}{c}{5178} & \multicolumn{1}{c}{5131} & \multicolumn{1}{c}{5131} & \multicolumn{1}{c}{5131} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} \\ 
Eff N Left & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{219} & \multicolumn{1}{c}{200} & \multicolumn{1}{c}{194} \\ 
Eff N Right & \multicolumn{1}{c}{147} & \multicolumn{1}{c}{157} & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{142} \\ 
BW Loc Poly & 8.970 & 9.368 & 8.739 & 8.530 \\ 
BW Bias & 13.860 & 15.029 & 14.752 & 13.035 \\ 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $ ^{***} $ p $ < $ .01; $ ^{**} $ p $ < $ .05; $ ^{*} $ p $ < $ .1. RD local linear estimates using \citet{calonico2019regression} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population, GDP per capita, number of seats in 2000, year, and dummy for Northeast region. \textit{N Left} and \textit{N Right} represent the total number of observation in the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

Note that without the covariates, the first stage overestimates the change in the primary treatment and underestimates the effect in the placebo regressions. When we add the covariates, the results improve and the point estimates are now one for the primary model and zero for the placebo regressions. These are the correct values. Moreover, Panel B of Table \ref{validitycheck} shows that the pre-treatment covariates remain unchanged around the thresholds. These tests reinforce the validity of our research design.

# Results
\label{sec:results}

To test the formal model's predictions, we present four sets of empirical results. First, we show that the alternative hypothesis, which emphasizes representation, competitiveness, and legislative productivity, fails to receive empirical support. Second, we confirm the bargaining costs hypothesis and find that the additional legislator also promotes more appointments of bureaucrats. This aligns with the councilors' descriptions of the strategies mayors employ to consolidate support from city councilors. Third, we show that municipalities with an extra city councilor approve more public goods laws in general and on education and health care. Finally, we show that all these pieces of evidence result in improvements in infant mortality and elementary school enrollment, two vital services for welfare.

## Council Size, Bargaining Costs, and Alternative Hypotheses
\label{sub:bargcosts}

Table \ref{tab:bargain} tests the first hypothesis and the alternative hypotheses about the nexus between city council size and bargaining costs. The formal model's central premise is that larger legislature sizes produce lower bargaining costs when the chances to elect politicians aligned with the mayor are sufficiently high. However, the results could stem from different mechanisms, such as minority and female representation, competition, and increased council-sponsored legislative productivity. In Panel A, we counter the alternative explanations, while in Panel B, we present evidence that corroborates our findings.

Panel A (Table \ref{tab:bargain}) tests hypotheses AH1 to AH4. AH1 and AH2 establish that better representation of under-served groups could improve welfare. However, columns (1) and (2) show that larger city councils do not increase the representation of females or non-whites. AH3 states that enhanced competition could improve responsiveness, increasing service provision. Looking at competition per seat in municipalities with larger city councils, Column (3) in Panel A shows a null effect of council size on the competition. Finally, AH4 states that council-sponsored legislation could be increasing in larger councils. Yet, Column (4) shows that it remains unchanged when council sizes increase.

\pagebreak

\begin{table}[!htbp] \centering 
  \caption{Political Effects of Increasing the Number of Legislators} 
  \label{tab:bargain} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
& \multicolumn{4}{c}{\textbf{Panel A: Representation, Competition, and Legislation Approval}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Num. Female}} & 
 \multicolumn{1}{c}{\small{Num. Non-white}} &
  \multicolumn{1}{c}{\small{Candidates}} & 
  \multicolumn{1}{c}{\small{Prop. Laws}} \\ 
   & \multicolumn{1}{c}{\small{councillors}} & 
 \multicolumn{1}{c}{\small{councillors}} &
  \multicolumn{1}{c}{\small{Per Seat}} & 
  \multicolumn{1}{c}{\small{Approved Council}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & 0.22 & 0.60 & -0.19 & -0.02 \\ 
  & (0.38) & (0.85) & (1.11) & (0.08) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5183} & \multicolumn{1}{c}{239} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{3424} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{158} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{269} \\ 
Eff N Left & \multicolumn{1}{c}{195} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{179} & \multicolumn{1}{c}{227} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{132} & \multicolumn{1}{c}{141} \\ 
BW Loc Poly & 8.623 & 2.958 & 8.144 & 10.976 \\ 
BW Bias & 13.565 & 4.922 & 12.018 & 16.708 \\[1.0ex] 
\midrule 
& \multicolumn{4}{c}{\textbf{Panel B: Partisan Alignment and Bureaucratic Appointments}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(5)}} 
 & \multicolumn{1}{c}{\small{(6)}} & & \\
 & \multicolumn{1}{c}{\small{Mayoral}} & 
 \multicolumn{1}{c}{\small{Num. Politically}} & & \\
 & \multicolumn{1}{c}{\small{Coalition Size}} & 
 \multicolumn{1}{c}{\small{Bureaucrats}} & 
  & \\[1ex] \cline{2-2} \cline{3-3}
 LATE & 0.91$^{*}$ & 104.49$^{*}$ & & \\ 
  & (0.50) & (61.53) & & \\ 
    \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5179} & \multicolumn{1}{c}{15536} & & \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{1027} & & \\ 
Eff N Left & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{372} & & \\ 
Eff N Right & \multicolumn{1}{c}{162} & \multicolumn{1}{c}{343} & & \\ 
BW Loc Poly & 10.025 & 6.242 & & \\ 
BW Bias & 15.889 & 10.243 & & \\ 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

Panel B (Table \ref{tab:bargain}) illustrates the mechanism put forward in the formal model. Column (5) shows a 91\% chance of the extra legislator belonging to the mayoral coalition. The model hypothesis stated that the probability of electing a city councilor allied with the mayor is sufficiently high. However, sufficiently high can be ambiguous, and it depends on the intensity of the political disagreement and the current legislature size. In any case, 91\% chance is a considerably large quantity, and the extra city councilor is going to the mayoral coalition nine in every ten legislature size increases. This strengthens the political position of the mayor and diminishes bargaining costs.\footnote{The effect of city council size on the number of mayor-aligned politicians stems from the rounding rule adopted by the proportional representation system in Brazil. The rounding rule favors the coalitions with more votes. Most of the time, the coalition of the elected mayor is the one that receives more votes.} Column (6) demonstrates that adding one councilor increases appointed employees by around 104 extra employees per additional city councilor. When negotiating with city councilors, these employees are a great bargaining chip, and they are also associated with better service provision.

## Approved Legislation and City Councilor's Survey
\label{sub:survey}

We built two novel datasets to provide evidence for our mechanism. First, we ran a survey with former city councilors, asking how the mayor consolidate its coalition support within the city council. Second, we collected 346,553 laws approved in 63 municipalities within ten thousand inhabitants away from the city council size thresholds. We hand-coded 3,466 laws into four categories: public goods, oversight, education and health care-related legislation, and other legislative duties. After hand-coding these laws, we used a Supporting Vector Machine (SVM) algorithm to classify the remainder laws. From all of the algorithms we tested, SVM achieved the highest performance.

\begin{figure}[htb!]
  \begin{center}
  \caption{Mayoral Instruments for Securing Councilors' Support}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=6in,angle=0]{../figures/strategySupportMayors.pdf}}
      \bigskip
      \label{survelec}
    \end{minipage}
  \end{center}
\end{figure}

This paper assumes that city councilors want to extract rents and to get reelected. The rents come from continuing in the repeated game, and the partisanship discounts that benefit. Figure \ref{survelec} shows that 65.9 percent of councilors mentioned that mayors use job appointments to consolidate their coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed that mayors use personal favors, such as shifting resources to places where the councilors have electoral support, to solidify their governing coalition. Granting favors works as rent transfers for councilors. Job appointments are also convenient for the mayors because patronage in Brazil has positive results in terms of welfare \citep{toral2019benefits}.

\begin{figure}[htb!]
  \begin{center}
  \caption{Legislation Approval Patterns in the Municipalities Closer to the City Council Size Thresholds}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=6in,angle=0]{../figures/lawsApproved.pdf}}
      \bigskip
      \label{LEGAPPR}
    \end{minipage}
  \end{center}
\end{figure}

We have established in Table \ref{tab:bargain} column (4) that the quantity of laws councilors approve remain constant. However, this tells us little about the legislation composition in municipalities with larger council sizes. In Figure \ref{LEGAPPR}, we show that the per capita number of laws related to local public goods provision increases in towns that receive the extra legislator, vis-à-vis baseline cities that did not get an additional councilor. One additional city councilor increased the per capita amount of laws on public goods by 15.0 percent. Also, the number of laws that focus on healthcare and education increased by 19.7 percent. Looking specifically into healthcare and education legislation, the proportion of laws discussing public goods increased by 15.7 percent, showing that policy proposals by both the mayor and the councilors shift toward local and citywide public goods.\footnote{Public goods here are not restricted to citywide goods. Any service that benefits citizens, regardless of whether they are targeted to citizens or not, is considered public goods. For example, building a health clinic in a particular district can virtually exclude the citizens that are not living in the neighborhood. However, we am still considering these as public goods, albeit targeted at a specific locality.}

## The Effect of Legislature Size on Welfare
\label{sub:effects}

We now study the impact of legislature size on education and health care. We use four indicators in this paper: two in education and two in health care.

\begin{table}[!htbp] \centering 
  \caption{Legislature Size and Welfare} 
  \label{tab:outcomes} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
& \multicolumn{4}{c}{\textbf{Council Size and Welfare -- Main Estimates}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Infant}} & 
 \multicolumn{1}{c}{\small{Postneonatal}} &
  \multicolumn{1}{c}{\small{School}} & 
  \multicolumn{1}{c}{\small{Education}} \\ 
   & \multicolumn{1}{c}{\small{Mortality}} & 
 \multicolumn{1}{c}{\small{Mortality}} &
  \multicolumn{1}{c}{\small{Enrollment}} & 
  \multicolumn{1}{c}{\small{Quality}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & -2.01$^{**}$ & -0.90$^{*}$ & 2.52$^{***}$ & -0.03 \\ 
  & (0.79) & (0.48) & (0.80) & (0.13) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{12299} & \multicolumn{1}{c}{5441} & \multicolumn{1}{c}{10156} & \multicolumn{1}{c}{8597} \\ 
N Right & \multicolumn{1}{c}{1029} & \multicolumn{1}{c}{672} & \multicolumn{1}{c}{686} & \multicolumn{1}{c}{669} \\ 
Eff N Left & \multicolumn{1}{c}{501} & \multicolumn{1}{c}{324} & \multicolumn{1}{c}{206} & \multicolumn{1}{c}{409} \\ 
Eff N Right & \multicolumn{1}{c}{381} & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{291} \\ 
BW Loc Poly & 7.527 & 7.564 & 5.425 & 9.05 \\ 
BW Bias & 11.95 & 11.01 & 10.58 & 15.107 \\ [1.0ex] 
\midrule 
& \multicolumn{4}{c}{\textbf{Council Size and Welfare -- Placebo Estimates}} \\ \cline{2-5}
& \multicolumn{1}{c}{\small{(5)}} 
 & \multicolumn{1}{c}{\small{(6)}} 
 & \multicolumn{1}{c}{\small{(7)}} 
 & \multicolumn{1}{c}{\small{(8)}} \\
& \multicolumn{1}{c}{\small{Infant}} & 
 \multicolumn{1}{c}{\small{Postneonatal}} &
  \multicolumn{1}{c}{\small{School}} & 
  \multicolumn{1}{c}{\small{Education}} \\ 
   & \multicolumn{1}{c}{\small{Mortality}} & 
 \multicolumn{1}{c}{\small{Mortality}} &
  \multicolumn{1}{c}{\small{Enrollment}} & 
  \multicolumn{1}{c}{\small{Quality}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & -1.14 & -0.19 & 0.44 & -0.10 \\ 
  & (1.05) & (0.22) & (0.61) & (0.11) \\ 
    \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{10613} & \multicolumn{1}{c}{17236} & \multicolumn{1}{c}{9040} & \multicolumn{1}{c}{7528} \\ 
N Right & \multicolumn{1}{c}{2715} & \multicolumn{1}{c}{3624} & \multicolumn{1}{c}{1802} & \multicolumn{1}{c}{1738} \\ 
Eff N Left & \multicolumn{1}{c}{1210} & \multicolumn{1}{c}{780} & \multicolumn{1}{c}{1210} & \multicolumn{1}{c}{1255} \\ 
Eff N Right & \multicolumn{1}{c}{890} & \multicolumn{1}{c}{664} & \multicolumn{1}{c}{728} & \multicolumn{1}{c}{752} \\ 
BW Loc Poly & 4.838 & 2.807 & 6.364 & 6.873 \\ 
BW Bias & 7.143 & 8.041 & 9.68 & 10.291 \\
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

Table \ref{tab:outcomes} shows that increasing the city council by one legislator generates welfare improvement in municipalities. In health care, increasing the legislature size decreases infant mortality in 2.01 deaths per 1,000 children born. This impact is significant in absolute terms and represents an effect of 0.14 in standard deviations. Moreover, postneonatal mortality decreases by 0.90 casualties per 1,000 births, which remains around the 0.14 standard deviations improvement as infant mortality. Infant mortality is a primary outcome for welfare, and it correlates with several other indicators of well-being \citep{world2005health}. The fact that infant mortality improved in the Brazilian case shows that the increased mayor's support mechanism results in sizable welfare effects. As expected, placebo regressions were statistically insignificant.

In terms of education, with the addition of a city councilor, enrollment in elementary schools increased by 2.58 children on average per school. This represents a change of around 0.20 standard deviations. Besides the influence of the mayoral support, enrollment can also directly impact the councilor's representation: Councilors can pressure the bureaucracy to accommodate a few extra students. With support from the mayor, they can help voters get access to schools, improving the voter's welfare. The education quality remains unchanged, but with more students in the classroom, it could potentially decrease. As expected, the placebo regressions are insignificant.

# Discussion

This paper shows that partisanship considerably decreases bargaining costs, improving welfare. The crucial actors in providing public goods are the mayors, but they cannot do it alone. Legislatures are a solid check to the mayor's policy. Bargaining with them in a country plagued by corruption can be costly \citep{bertholini2017price}. Therefore, an extra ally in the city council may help ease the bargaining and improve welfare.

Increasing the council size results in gains in elementary school enrollment and in lower infant mortality. Education quality remains unchanged, but increasing enrollment without affecting quality should be interpreted as positive. The analysis of the approved legislation shows that the frequency of laws on local public goods increases in municipalities above the council size thresholds. Finally, the city council survey demonstrates that councilors and mayors agree about appointing more public servants to the public administration.

\citet{Ames2001} argues that the Brazilian democracy is deadlocked because the Executive has its hands tied by a legislature that demand constantly provision of particularistic and private goods. Indeed, the most successful presidents after 1995 had to build multiparty coalitions by negotiating with parties \citep{limongi1998bases, raile2011executive}. Despite evidence that side payments cemented these coalitions, many social and economic policies have been implemented. For instance, education and health care notably improved during the Cardoso and Lula governments \citep{acemoglu2012nations}. Of the presidents that defied the coalition system, Rousseff was impeached in her second term. Bolsonaro can be considered one of the most ineffective presidents in Brazilian history.

Municipalities work very similarly in terms of the incentives for legislative-executive bargaining. The stakes are lower, and the negotiations involve only one legislative branch. Still, in a system such as Brazil's, with more than thirty-five registered parties, mayors face substantial challenges to the support for their agendas. Having extra help in the council can tilt the bargaining towards cheaper governing, thus improving public goods provision.

Moreover, we do not argue that these benefits result from the absence of corruption or clientelism. Both the formal model and empirical evidence provided by \citet{britto2020corruption} show that corruption increases with legislature size. As the model shows, city councilors receive a side transfer from mayors to support their agenda. Mayors also divide the resources saved in the bargain into public goods and rents. The rate for the division depends on the marginal changes of the rents and public goods, but these two components receive the extra resources. In this case, the mayor's rents may be primarily corruption, and the councilor's transfers can be clientelistic, aimed at improving the councilor's reelection chances.

Our findings have implications that extend well beyond Brazil. As several countries have local city councils and multiparty systems, the dynamics we expose here may explain how fragmented legislatives are not synonymous with policy paralysis. This runs in contrast with well-established theories that point out that multipartism may be a peril to democratic regimes \citep{mainwaring1993presidentialism, stepan1993constitutional}. Additionally, our findings may help us understand why other countries in Latin America and Asia can generate effective governance despite having strong executives and large coalitions \citet{pereira2012surprising}.

This paper opens two questions for future analysis. First, if the mechanism we suggest is indeed at work, more support for the Executive means better public goods. Therefore, one may ask what should be the city council's checking prerogatives, as the they significantly reduce the provision of public services. The evidence so far is inconclusive, and \citet{poulsen2018political} show that checks for the mayors may also improve service provision. It may be the relationship between rents and public goods is non-linear, making mayors revert to only rents past a certain threshold. Moreover, it could be that, at some point, legislative support is ineffective or even harmful to democracies, as it may trigger an intense form of elite capture. This question needs further investigation, and our results suggest that mayors with more support improve welfare. Second, it remains uncertain how other legislative features affect welfare. For example, city councils may vary in size, monetary compensation for councilors, committee structure, and internal power structure. These components may interact with each other, and understanding how size plays with other features would improve our knowledge about city councils size affect welfare. Both these questions are crucial for institution design and welfare in developing democracies.

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}