### 2022-04-15 article/article.Rmd
add large text revision

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: 
date: \today
abstract: "How does legislature size impact public service provision? Despite the importance of institutional design for democratic governance, the effect of legislative features on citizen welfare remains little understood. In this paper, we use a formal model to show that increasing legislature size improves public goods delivery. We argue that changes in bargaining costs depend on whether additional legislators share the executive's party affiliation: More opposition members reduce the executive's public goods provision, while more government-aligned members increase it. We test this theory by exploiting sharp discontinuities in city council size in Brazil. We show that an additional city councilor has a 91% chance of belonging to the mayoral coalition, and this significantly improves primary school enrollment and infant mortality measures. To explore possible mechanisms, we surveyed former city councilors and analyzed 346,553 bills proposed between 2005 and 2008. This paper has implications for the design of representative institutions."
abstractspacing: double
keywords: Legislature Size; Legislative Bargaining; Local Public Goods Provision; Regression Discontinuity Design; Brazil
jelcodes: H41; H80; I31; I38
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
# Knitr options
knitr::opts_chunk$set(fig.pos = "H") # holds figure position
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

# Introduction
\label{sec:intro}

Legislative institutions are crucial for citizen welfare and public service
provision in democratic countries. Historically, the division of power and the
birth of representative governments placed legislatures at the center of the
political stage. In modern democracies, most countries have assemblies at all
levels of government. These institutions regulate the quantity and quality of
legislation, check and veto other political powers, and influence policymaking.
Moreover, legislatures also determine taxation, government spending and oversee
public programs \citep{auriol2012optimal, north1989constitutions,
poulsen2018political, stasavage2011states, Weingastetal1981}.

Regardless of eventual differences in electoral rules, social composition, and
internal procedures, all legislatures share one common feature: They are
collective bodies. As such, the number of elected representatives has important
consequences for the whole polity. Legislature size is known to have a
significant effect on collective action strategies, the bargaining power
between state branches, government spending, and representation
\citep{crain1979cost, freire2021legislature, gerring2020population,
PrimoSnyder2008, rogers2002free, Weingastetal1981}. However, how legislature
size impacts social welfare remains less understood. On the one hand, enlarged
legislatures may reduce political particularism and increase overall government
spending, which can lead to more public goods \citep{gerring2020population,
GilliganMatsusaka2001, Weingastetal1981}. On the other hand, adding more
representatives to a legislature may exacerbate collective action problems,
thus hindering service provision \citep{gottlieb2019countervailing,
tsebelis2002veto}. Therefore, what is the effect of larger legislatures on
public service delivery?

In this paper, we propose a new theory based on executive leadership and party
politics to explain how city council size impacts local public goods provision.
First, in mayor-council municipalities, mayors are the main responsible for
public service decisions and have the power to boost public investment similar
to presidents and governors. Second, partisanship plays a key role in city
politics, even in countries in which ideological attachment is relatively weak.
Party attachment structures political views \citep{samuels2014power}, and
local-level elites usually belong to different factions and parties. They fight
for power in the municipality by controlling various parties and forming
distinct pre-electoral coalitions.\footnote{Elite political competition in the
local level has also been linked to violence and conflict in Brazil and
elsewhere \citep{hoelscher2015politics, mignozzetti2018predatory}.} In many
cities, family ties may be substitutes for the ideological component, but
influential local leaders usually sort themselves into different political
groups \citep{ferraz2020political, nunes1977coronelismo}.

This paper incorporates partisanship and mayoral first-move advantage to a
\citet{baron1989bargaining} bargaining model. Mayors make the initial policy
proposal, which consists of investments in public goods provision, rents for
personal gains, and a vector of side-transfers for councilors. If mayors fail
to secure majoritarian support from councilors, a reversal stage occurs, in
which the decision leaves the mayor's office and councilors propose the public
investments. In this mayor-council bargaining process, costs come from the
side-transfers required to secure a majority among councilors. We show that in
a model without partisan concerns, bargaining costs always increase with larger
city councils, which in turn lower public goods provision and rents. However,
partisanship decreases bargaining costs when the chances of electing councilors
aligned with the mayor are sufficiently high. Lower bargaining costs release
funds for mayors to invest in public goods provision. In this case, a larger
city council may have more partisan councilors aligned with the mayor, reducing
the mayor's bargaining costs and increasing the equilibrium public goods
provision.

We test these theoretical predictions by exploiting exogenous variation in city
council size in Brazil. In 2004, the judiciary reinterpreted the Brazilian
Constitution and unexpectedly changed the number of councilors in several
municipalities. Until then, cities could freely determine their legislature
size, but in March 2004 the Supreme Electoral Court created population
thresholds and added one councilor to each set of 47,619 inhabitants. Around
these cutoffs, the decision represented an exogenous change in city council
size, allowing us to investigate the effects of increasing the number of local
legislators in Brazil.

We find that larger city councils changed the composition of mayoral coalitions
and the number of appointed bureaucrats in the municipality. The extra city
councilor had a 91 percent chance of belonging to the mayor's pre-electoral
coalition. With a larger coalition, mayors face lower bargaining costs and are
able to boost public investments with fewer constraints. Each additional
legislator also leads to 104 new appointed bureaucrats, and research shows that
political appointees civil servants improve public service provision and
enhance accountability in Brazil \citep{toral2019benefits}.

To provide further evidence for our mechanisms, we surveyed 174 former
councilors who served during the 2005-2008 term, the period we analyse in this
paper. They confirm that mayors indeed use bureaucratic appointments and
political favors to secure legislative support. As our formal model shows,
bargaining with councilors requires offering transfers to the representatives.
By asking councilors what they would like to provide to voters, we find that
councilors mainly demand job appointments, as these increase their reelection
prospects and improve municipal welfare. These findings are in line with
quantitative results from a novel dataset we built with 346,553 laws passed in
63 municipalities within ten thousand inhabitants away from the cutoffs. We see
that legislation in municipalities both on the left- and on the right-hand side
of the cutoffs mention public goods provision. However, municipalities with
larger city councils have 15 percent more proposals about public goods
provision, which is consistent with our argument.

Finally, we show that the increase in legislature size significantly impacts
municipal education and health care. Increasing the city council size by one
legislator lowers infant mortality by 2.01 children per one thousand infants
born and reduces postneonatal mortality by 0.90 children per one thousand
infants that survived their first 28 days. Larger councils also increase
enrollment by 2.58 children in elementary education classrooms without reducing
school quality. These services are crucial for citizen welfare and result from
eased bargaining between the mayor and the city council.

This paper contributes to four strands of the literature. First, this paper
dialogues with a large literature on multilateral bargaining. Repeated models
of multilateral bargaining started with \citet{baron1989bargaining}, who
extended the bilateral bargaining game proposed by
\citet{rubinstein1982perfect} to more than two players. Several papers followed
\citet{baron1989bargaining} and included endogenous recognition power, the size
of the required majority, and partisanship in the model
\citep{agranov2020persistence, baranski2020determinants,
calvert2005legislative, choate2019partisan}.\footnote{See
\citet{eraslan2019legislative} for a comprehensive review of multilateral
bargaining and \citet{ray2015coalition} for a review on coalition formation.}
In this paper, we extend the \citet{baron1989bargaining} model under closed
rules and add the mayor's first move and political partisanship. We consider
partisanship as a zero-sum effect that is positive for politicians aligned with
the mayor and negative otherwise. This paper investigates the often overlooked
effect of legislature size on the bargaining equilibrium strategies.

Second, this paper contributes to the literature on the effects of larger
legislatures in service provision. Following the seminal contribution by
\citet{Weingastetal1981}, the "law of $1/n$" posits that larger legislatures
increase the legislative pressure for providing targeted public goods to local
supporters at the expense of the entire polity \citep{freire2021legislature}.
This question has become more relevant since legislature size has increased
consistently over time \citep{gerring2020population}. The corollary of the "law
of $1/n$" is that increasing legislature size promotes excessive public
spending, regardless of the partisan makeup of the legislature
\citep{GilliganMatsusaka2001}.\footnote{\citet{PrimoSnyder2008} show that
legislature size may result in positive or negative expenditures, depending on
factors such as the degree that local public goods can exclude non-locals or
deadweight costs associated with taxation.} In the US metaphor, this would
explain significant infrastructure spending in smaller districts. However, the
"law of $1/n$" literature rarely mentions the relationship between public goods
provision and welfare. This paper is one of the first to study how the number
of local legislators affects welfare. Our model provides a theoretical
expectation based on the chances of electing a politician aligned with the
mayor. The empirical estimates confirm our theoretical predictions, with the
additional fact that legislature size has a null effect on expenditures in
Brazil.

Third, our article also contributes to the growing literature on municipal
partisan politics. Most papers in this field estimate the effects of the
mayor's party on political outcomes. For instance, \citet{de2016mayoral} and
\citet{gerber2011mayors} find differences in governance when comparing
Democratic and Republican mayors. Specifically about Brazil,
\citet{gouvea2021partisanship} shows that social policy expenditure increase
with left-wing mayors. \citet{frey2021legislature}, using discontinuities in
city council size after the 2009 constitutional amendment that reverted the
thresholds we study here, shows that changes in council size lower the support
for candidates backed by the municipal mayors. Our paper reinforces the view
that partisanship matters locally, even in countries with weak ideological
linkage.

Finally, this paper contributes to the literature on the causes and effects of
legislature sizes. While most papers on legislature size analyse the trade-off
between representation and efficiency \citep{frederick2008people,
jacobs2015explaining, muzzio1989size, Stigler1976}, recent articles have
explored the effects of legislature size on lawmaking and partisanship in
further detail. \citet{kirkland2014chamber} shows that larger legislatures make
it harder for politicians to know and collaborate with each other, so they rely
more heavily on party cues and committee structures. In considering the causes
of changes in legislature sizes, \citet{jacobs2015explaining} show that
population shifts increase and economic crises decrease the number of
legislators. \citet{marland2019fewer} questions the economic crises mechanism,
showing that political elites shrink legislatures mostly to signal to the
public the priority of their austerity agenda. In this article, we show that
council size affects welfare by changing the partisan composition of
the legislatures.

# A Theory of Legislature Size, Legislative Bargaining, and Welfare
\label{sec:theory}

Institutional design has a major impact on long-term welfare
\citep{acemoglu2012nations, buenodemesquita2005logic, north1989constitutions}
and democratic consolidation \citep{gandhi2019institutional,
lijphart2018institutional}. One particular problem that all polities face is
how to determine the optimal size of their legislative bodies. The literature
on this topic is extensive, and modern attempts to provide a solution to the
legislature size problem started in the early eighteenth century with
the Federalists versus Anti-Federalists debates \citep{muzzio1989size,
Stigler1976}.

The literature on legislature size highlights that the number of legislators
causes tension between representation and efficiency
\citep{frederick2008people, jacobs2015explaining, muzzio1989size,
rush2013conclusion, Stigler1976, taagepera1972size}. Larger legislatures
improve representation as both the election thresholds and the ratio of
population-per-politician decrease, which facilitate the electoral success of
ethnic and political minorities \citep{allenrepresentation,
gerring2020population}. However, better representation is often countered by
the decline in efficiency in larger legislatures \citep{Weingastetal1981}. The
enlarged size makes it harder for members to coordinate and collaborate without
partisanship and legislative structures \citep{kirkland2014chamber,
rush2013conclusion}. It also increases the potential number of veto players,
decreasing policymaking efficiency \citep{tsebelis2002veto}. Balancing these
two incentives is crucial for the optimal legislative composition
\citep{auriol2012optimal, taagepera1972size}.

This literature has often overlooked two important questions. First, does
legislature size matter in contexts where legislative bodies are not the
pivotal decision-maker? And second, what is the ultimate effect of legislature
size on welfare? These gaps are likely generated by the predominance of the
"law of 1/n" scholarship in political economy debates, which focus mainly on
the relationship between legislature size and overspending. Since
\citet{Weingastetal1981}, 30 papers have systematically tested the effects of
legislature size on government expenditures \citep{freire2021legislature}. Of
these papers, only \citet{lewis2019legislature} investigates its effects on
welfare, showing that larger legislatures have a negative welfare
effect.\footnote{\citet{koppel2004public} also shows, using a repeated game
where the proposal prerogative rotates in a legislature, that larger
legislatures increase the chances of non-cooperative outcomes, lowering public
goods provision.} \citet{frey2021legislature} analyzes the Brazilian case after
the 2009 legislature size change and finds a null effect in service provision
from 2012 to 2016. Yet, \citet{GilliganMatsusaka2001} finds that in the US
legislature size is associated primarily with higher spending on geographically
concentrated benefits, such as education and highways. These results suggest
that the legislature size and welfare nexus needs a dedicated theoretical and
empirical assessment.

Despite the importance of legislatures, the executive still dominates the
public debate on public service provision in most mayor-council (or
presidential) systems. For instance, while in 2021 the New York Times cited
Mayor Bill De Blasio 2,084 times, it mentioned the 49 seating city councilors
only 440 times in the same year. When we expand the search for the past five
years, we see that De Blasio was mentioned 8,978 while the 49 seating city
councilors were mentioned 2,046 times.\footnote{We computed these figures using
the Dow Jones Factiva Database in November 2021.} On average, the mayor was
cited four times more often than all the city councilors combined. This
executive dominance has implications for policymaking, as coalitions and
affinity with the executive are going to be crucial for service provision
\citep{buenodemesquita2005logic}.

## Model Setup

In a mayor-council setting, consider a strategic interaction between a mayor
($M$) and $N \geq 3$ city councilors indexed in $i \in \{1, 2, \cdots, N \}$.
The mayor has the prerogative of proposing a vector of policies on which city
councilors vote. The provision of public goods is then a bargain between the
mayor and the city council. Policy proposals are a combination of public goods
provision $g$, rents for the mayor $r$, and a vector of transfer for city
councilors \textbf{x}.

The city council votes the mayor's proposal, and if councilors accept it, the
proposal is implemented. Otherwise, a reversal policy takes place. Reversal
policy in this context means that the mayor's proposal failed to secure the
majority of council votes. The decision process then leaves the mayor's office
and is transferred to the council. The types of reversal policies are crucial
for our argument, as they change the relative strength of the council when
legislators bargain with the mayor. We investigate two types of reversal
policies: A baseline reversal with no parties and a hybrid mechanism with
partisan and non-partisan concerns. The differences between these two
mechanisms are that the non-partisan councilors only care about rents, while
partisan councilors care about both rents and policy.

The mayor must convince at least half of the city councilors to support her
policy proposals. Similar to \citet{buenodemesquita2005logic}, we assume that
the selectorate, that is, the individuals who have a say in the mayor's policy
proposals, are the $N$ city councilors. The winning coalition, the minimum
number of members that must be convinced to implement the policy, is equal to
half of the selectorate ($N/2$). City councilors are motivated by policy
concerns $p$ and transfers received from the mayor or the councilor nominated
as the proposer \textbf{x}. We leave the definition of these transfers open, as
this allows us to analyze how mayors combine transfers with other incentives to
pressure city councilors. Exemples of transfers are granting electoral favors,
funding clientelistic brokers, spending in areas that the councilor has
political or personal interests, providing portfolio (municipal offices) and
public jobs within the municipal bureaucracy, or even giving campaign
contributions and bribes. The use of transfers to influence the councilors'
choices generate a bargaining cost of $C_G$. Bargaining costs vary according to
the reversal mechanism and the size of the legislature.

Finally, after learning the bargaining costs, the mayor provides a level of
public goods provision $g$ and pockets rents $r$. Public goods provision helps
the mayor to get reelected. Rents are for the mayor's direct consumption and do
not contribute to the mayor's electoral success.\footnote{This model, similar
to \citet{cowen1994rent} and \citet{buenodemesquita2005logic}, combines public
goods provision and rent extraction. For a dynamic bargaining model that
incorporates transfers to districts and public goods provision, see
\citet{battaglini2012legislative}.} This makes the mayor's expected utility a
sum of the gains from rents and the benefits from reelection. Both the
utilities from rents $u(.)$ and the probability of reelection $\pi(.)$ are
concave functions, meaning that more rents or public goods increase the utility
at a decreasing rate. The probability of reelection is multiplied by the
benefit from holding office $B_M > 0$. This benefit captures the tangible and
intangible gains that the mayor receives from holding the public office. We
assume that the benefits are high enough to rule out an equilibrium with only
rents and transfers, which is equivalent to saying that mayors strongly prefer
to be reelected. If the mayor's policy proposal is approved, then the expected
utility for the mayor is:

\[
\mathbb{E}U(r, g) \ = \ u(r) + B_M\pi(g)
\]

The policy choices of the mayor are subject to municipal budget constraints.
The municipality has $R > 0$ resources and cannot run debts, meaning that the
budget must be balanced. We also assume that there are enough resources for the
mayor to govern.

Let $C_G$ be the bargaining costs. The budget balance constraint requires that
the offers the mayor makes to the city councilors must satisfy the following
inequality:

\[
r + g + C_G \ \leq \ R
\]

The expected utility for the city councilors depends on the type of reversal
policy. In the baseline non-partisan reversal, when the council rejects the
mayor's proposal, a reversal stage starts with resources diminishing by a
factor of $\delta \in (0,1)$ and the random selection of one councilor. The
selected councilor becomes the next proposer. If the new proposal is accepted,
it is implemented. If councilors reject the offer again, the budget shrinks by
$\delta$ and another city councilor is nominated to propose. The process
repeats until a proposal is finally accepted.

We now consider a situation that combines both partisan and non-partisan
legislature concerns. Each councilor has a party affiliation. Party
affiliations are mutually exclusive, and if a city councilor is aligned with
the mayor's party, then she belongs to the government coalition $G \subset \{1,
2, \cdots, N \}$. Otherwise, the city councilor belongs to the opposition $O =
\{1, 2, \cdots, N \} \setminus G$. The mayor's policy generates a political
value of $p > 0$ for the councilors, and whether it adds or subtracts from the
councilor's utility depends on the party affiliation of the
councilor.\footnote{One example of partisan benefits is that, as the mayor
shares partisan ties with members of the government coalition, a fraction of
the public goods provision can be claimed by councilors as their contribution.
We propose an extension in the Online Supplementary Materials, which assumes $p
= \phi g$, where $\phi$ is the fraction of the public goods provision that can
be credit claimed by the city councilor aligned with the mayor.} If councilors
reject the mayor's proposal, no public goods are provided ($p=0$). Moreover, in
equilibrium, we assume that city councilors would prefer to receive transfers
instead of providing public goods.\footnote{In the Online Supplementary
Materials, we consider a version of the model in which public goods provision
depends on the party affiliation of the proposer so that it affects choices in
any potential history of the game. In equilibrium, the level of public goods
provision is even higher than in the model presented here, as city councilors
prefer to approve the mayor's proposal than run the risk of an opposition
member becoming the proposer.}

The timeline of the strategic interaction is as follows:

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, x)_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts the proposal, the policy is implemented and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) The budget is discounted by a factor $\delta \in (0,1)$, and one councilor is randomly selected to make an offer. If the offer is accepted, it is implemented. If rejected, the reversal policy stage restarts.\footnote{In section A.1 of the Supplementary Materials, we solve three reversal mechanisms. We include the two mechanisms we present here, plus a third mechanism with only political concerns.}
\end{enumerate}

The solution concept we use in this model is the subgame perfect Nash
equilibrium. This game requires that the strategies follow a Nash equilibrium
in each subgame. We find the solution using backward induction. In the model
with infinitely repeated proposals, any strategy can be shown to be a Nash
equilibrium. Thus, a natural assumption is to extend the equilibrium concept to
require the equilibrium to be stationary. A stationary subgame perfect
equilibrium requires that, at each given point of the game, if a politician
accepts an offer at time $k+1$, she should take the same offer at time $k$
\citep{rubinstein1982perfect}. Stationarity gives us a method to find a
proposal that would be accepted at any stage of the game. Hence, we can
characterize a sequence of offers for each point $k$ in time. For the mayor,
the optimal offer is at $k=0$, representing a no delay in policy
implementation. We assume that players have no dominated strategies, even when
there is insufficient support for the policy.

For the solution to be of general interest, we solve the bargaining costs for
the expected number of opposition and government-aligned city councilors. This
is because the solution for each fixed number of opposition and government
councilors does not help us understand the overall relationship between council
size and bargaining costs.

## The Mayor's Decision Stage

Solving the game by backward induction requires us to start with the
councilors' strategies. However, to conserve space, assume that we solved the
game for the councilors and found the equilibrium bargaining costs $C_G(N)$
when there are $N$ city councilors. We now derive the optimal rents ($r$) and
public goods provision ($g$) proposed by the mayor. The mayor benefits from
public goods provision, as it increases the chances of her reelection. However,
she prefers to invest as minimum as necessary in public goods and extract the
remaining resources as political rents. The mayor's objective is to maximize
her expected utility, subject to the municipal budget constraint.

\[
\begin{aligned}
\max_{r, g} \quad & u(r) + B_M\pi(g)\\
\textrm{s.t.} \quad & r + g + C_G(N) \leq R
\end{aligned}
\]

In equilibrium, the budget binds, and the first-order condition for optimal
public goods provision indicates that the marginal costs of providing public
goods is equal to the marginal benefits of reelection:

\[
u'(R - g - C_G(N)) = B_M\pi'(g)
\]

The optimal level of public goods provision equates the marginal gains accrued
from rents with the marginal benefits from reelection.

\begin{proposition}\label{prop1}
Public goods provision increases with legislature size when the bargaining costs decrease with legislature size.
\end{proposition}

All proofs are in section A.1 of the Supplementary Materials. The proof's
intuition involves finding what happens to the mayor's marginal utility when we
increase both $g$ and $N$. As $N$ increases discretely on a complete lattice,
we use monotone comparative statics to derive the conditions for increasing
differences \citep{milgrom1994monotone, ashworth2006monotone}. Increasing
differences mean that higher $N$ leads to higher equilibrium $g$.

Proposition \ref{prop1} provides our first empirically testable hypothesis:
\textit{if bargaining costs decrease when legislature size increases,
then public goods provision increases when legislature size increases}. Now
we study each reversal mechanism to determine the condition under which the
bargaining costs decrease when legislature size increases.

## Baseline: Non-Partisan Legislative Bargaining

In the non-partisan reversal mechanism, we assume that when the council rejects
the mayor's office, a reversal stage starts with the random selection of one
councilor. This formulation is similar to the closed rules proposal in
\citet{baron1989bargaining}. To find a stationary subgame perfect Nash
equilibrium, suppose that there were $k-1$ rejections, and the game is at the
$k$-th stage. A councilor accepts the proposer's offer if, and only if, to
accept the offer at $k$ is better than or equal to wait until the next stage
$k+1$. If the offer is $x_i$, then:

\[
x_i \geq \dfrac{1}{N}\left[ \delta^{k+1}R - \dfrac{N}{2}x_i \right] + \left[ 1 - \dfrac{1}{N} \right]\left[\dfrac{1}{2} x_i \right]
\]

We place the offer on the left-hand side. There are two components on the
right-hand side. The first is the amount that the councilor $i$ gets when she
is the proposer. It is equal to the chance that she is recognized as the
proposer times the budget in the next round minus the offers she makes to
convince half of the councilors. The second part represents the gains if the
councilor rejects the current received offer but still gets an offer in the
following round. It is equal to the chance that she is not recognized as the
proposer times the chance that she receives a transfer times the transfer
amount. Note that the offer the councilor makes as the proposer is the same as
the offer she wants to receive. This is because the city councilors are
exchangeable. The solution is symmetric for all councilors receiving an offer
(this means that, without loss of generality, we could have dropped the $i$ in
the solution). After some algebra, the offer $x_i$ must be greater than or
equal to:

\[
x_i \geq \dfrac{2 \delta^{k+1}R}{2N + 1} \equiv \underline{x}(k, N)
\]

The proposer always offers the minimum required to get the proposal approved.
In this case, the offer at any given stage $k$ is going to be equal to
$\underline{x}(k, N)$.

Proceeding backwards, at the mayor's proposal stage $k=0$ the mayor is going to
offer $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the
councilors. In this context, the bargaining costs for the mayor are:

\[
C_G(N) \ = \ \dfrac{N}{2}\left[\dfrac{2 \delta R }{2N + 1}\right] \ = \ \dfrac{\delta R N}{2N + 1}
\]

\begin{proposition}
In the baseline non-partisan reversal, bargaining costs always increase when legislature size increases.
\end{proposition}

The intuition for the results is that every additional councilor demands more
resources from the mayor. Thus, more councilors represent higher bargaining
costs. And if the bargaining costs increase with legislature size, then any
increase in public goods provision associated with legislature size has come
from a different mechanism. Below we demonstrate that partisanship is the key
to understand how this relationship works.

## Partisan and Non-Partisan Bargaining

In a reversal mechanism that incorporates partisan and non-partisan
motivations, a city councilor aligned with the government favors the mayor's
offer if:

\[
x_i \geq \dfrac{2\delta R}{2N+1} - p
\]

An opposition politician, on the other hand, favors the mayor's offer if, and
only if:

\[
x_i \geq \dfrac{2\delta R}{2N+1} + p
\]

We decompose these costs into two components. First, the costs in terms of
rents. Second, the costs (or benefits) from political alignment. In this
context, the bargaining costs depend on the mayor's offers for councilors in
her coalition versus the mayor's proposals for councilors in the opposition. As
the ex-ante chance of an opposition member being elected is equal to
$1-\gamma$, after taking the weighted averages these costs become:

\[
C_G(N) \ = \ \dfrac{N}{2}\left( \dfrac{2\delta R}{2N+1} - p \right) + \left(\dfrac{N}{2} - \gamma N \right)2p
\]

\begin{proposition}
In the reversal mechanism that incorporates partisanship, if $\gamma \geq \dfrac{1}{p}\left[ \dfrac{1}{(2N+1)(2N+3)} \right] \equiv \underline{\gamma}$, then bargaining costs decrease when legislature size increases.
\end{proposition}

The intuition for the proof is that as partisanship influences the costs and
benefits of legislators, more government-aligned councilors make it easier for
the mayor to govern. When the partisan benefit is higher than the additional
cost that a legislator generates, adding a councilor will benefit mayors if the
councilor has a higher chance of sharing the same party as the mayor.
Therefore, a higher probability of electing a legislator aligned with the mayor
reverts into lower bargaining costs.

Also, the threshold $\underline{\gamma}$ that captures the probability of
electing a mayor-aligned policitian decreases along with the political value of
policy $p$ and the size of the legislature $N$. This means that intensive
partisanship and larger legislatures make it easier to satisfy the electoral
threshold.

## Main Hypotheses

Our model provides two empirically testable hypotheses. Consider a municipality
with a city council comprised of $N$ legislators. Then the city increases the
size of the council to $N+1$. Thus, in a partisan setting:

\begin{enumerate}[labelindent=*, 
style=multiline, leftmargin=*, 
label=H\arabic*.]
  \item Bargaining costs decrease with legislature size when the chances of electing a government-aligned legislator are sufficiently high.
  \item Public goods provision increases if bargaining costs decrease with legislature size.
\end{enumerate}

# Background on the Brazilian Case
\label{sec:background}

In 2004, Brazil comprised 5,560 municipalities. According to the Brazilian
Constitution, each city must provide health care, primary education,
transportation, and infrastructure to its citizens. Municipalities have the
authority to enact local laws and collect taxes on housing and services.
However, their room for fiscal manipulation is limited. On average, municipal
taxes account for only 5\% of the local budgets, while mandatory transfers from
state and federal governments correspond to the remaining 95\%. This reduces
the ability of municipal legislators to effectively increase local spending,
which makes improvements in local services dependent only on the composition of
their local investments.

With regards to their political organization, municipalities have one mayor and
one city council. Citizens choose their representatives by direct vote in the
same election. Mayors oversee the provision of public services and propose laws
and amendments to the tax code. Proposals submitted by the mayor are subject to
the approval of the city council. Councilors, in turn, have two primary duties.
First, they should discuss and vote on legislation put forward by the mayors or
by their peers. Second, they oversee the public budget and sanction mayors who
do not comply with fiscal legislation. Councilors and mayors also provide an
array of public and personal services for their constituents, such as helping
voters access public hospitals, facilitating school enrollment, and even paying
their voters' utility bills \citep{lopez2017legisladores,
nichter2011electoral}.

Until 2004, there was no specific regulation on the size of city councils. The
1988 Brazilian Constitution set only broad guidelines about the number of
council seats in each municipality, thus allowing cities to determine how many
legislators they should have in their local chambers. This led to severe
imbalances in the proportion of representatives per citizen across the country.
For instance, the Nova Russas municipality had 30,009 inhabitants and 21
councilors, accounting for one legislator per 1,429 citizens. In contrast, the
city of Sorocaba had only 15 councilors for 528,735 inhabitants, that is, one
legislator per 35,249 citizens. In 2003, this malapportionment problem gained
national visibility when Mira Estrela, a small municipality with only 2,651
residents, reduced its city council from 11 to 9 seats because of fiscal
considerations. The change motivated a legal dispute that reached the Brazilian
Supreme Court. The Court favored the city's decision and ruled that the size of
every local council should correspond to precise population thresholds.

Following the ruling of the Supreme Court, in March 2004 the Supreme Electoral
Court established a series of population cutoffs for all municipal
legislatures. Following the Brazilian Constitution, the Court specified that
cities with up to one million residents should have between 9 and 21
legislators. The number of seats was allocated as follows. Each town starts
with nine councilors, then adds one legislator for every 47,619 inhabitants
until their councils reach 21 members. This threshold implies that cities with
a population between 571,428 and one million should all have the same number of
legislators, that is, 21 of them. Conversely, cities between one and four
million inhabitants should have at least 33 councilors, and this number
increases up to 41 members following the previous formula of one additional
legislator per 47,619 inhabitants.

Brazil also adopts a unique formula to allocate seats, which combines the Hare
quota and the D'Hondt method and favors larger parties
\citep{nicolau2015sistemas}. This poses a major obstacle for small parties to
reach the electoral quorum, so they often join party coalitions to improve
their chances of winning legislative seats \citep[p. 68]{Ames2001}. In Brazil,
mayors are usually members of the largest coalitions
\citep{poulsen2018political}, thus there is an incentive for candidates to side
with the mayor before and after the election.

These factors make Brazil the ideal testing ground for our theory. Although an
endogenous decision motivated the Electoral Court ruling, the population
cutoffs created sharp discontinuities in the distribution of local council
seats. Cities close to the cutoffs had very similar characteristics, but those
just above the threshold gained a new legislator. Moreover, cities could not
self-select into any group, as their population estimates were calculated using
the 2003 projections by the Brazilian Census Bureau (IBGE). This change in
council size also happened only seven months before the election, making it
impossible for candidates to adjust their service provision strategies for the
2004 election. Therefore, the ruling allows us to study the effect of larger
legislatures on public service provision during the 2005-2008 term while
holding all else constant.

# Variables and Data Sources
\label{sec:sources}

We use four groups of city-level variables in our models. The first group
includes pre-treatment control variables that were not affected by the 2004
council size resolution. We add them to the estimates to improve the balance
between treated and non-treated units. The covariates are: 1) the number of
council seats in the previous term; 2) city population; 3) city GDP; 4) the
proportion of low-income families in each municipality. The data come from the
Supreme Electoral Court and the 2000 Brazilian census.

Second, we look into municipal data, city councilors' characteristics, approved
legislation, and an online survey with 174 former city councilors that served
during 2005--2008 term to investigate potencial political
mechanisms.\footnote{We also analyzed 108 structured interviews conducted by
\citet{lopez2017legisladores} at the Brazilian Economics Planning Institute
(IPEA) in 2009. The authors asked councilors about their daily work. The
interviews reinforced that the mayor's agenda dominates councilors' work and it
is partisan without strictly following party lines. Municipal politics is
divided into families or powerful groups that belong to different party
affiliations. Councilors from other parties gravitate around these groups,
hence the opposition and government-aligned logic.} The municipal data
contains: 1) the number of councilor's belonging to the mayor's pre-electoral
coalition; 2) the number of appointed bureaucrats; 2) the councilor's gender
and race; 3) the competitiveness of each city council seat; 4) the proportion
of laws proposed by the city council that was approved. All variables come from
the Supreme Electoral Court, the Senate legislative support Interlegis, and the
Brazilian Bureau of Statistics.

Third, we also collected 346,553 bills from cities within ten thousand
inhabitants from the 2004 population thresholds. We selected 63 out of 202
municipalities whose voting records were available online. We evaluate proposal
types, discriminating between local public goods, oversight, and other
legislative activity. Examples of the local public goods are investing in a
public health clinic, requesting school staff to accommodate an extra child,
renovating sewage, and fixing potholes on a given street.\footnote{Our theory
does not differentiate between clientelism and broad service provision, as it
is hard to disentangle differences in the local level legislation. A voter may
request a particular service, such as fixing a pothole. Attending this request
can be interpreted as a clientelistic practice. Still, the provision of the
service generates non-excludable benefits to the adjacent neighborhood.}
Oversight bills are accountability and transparency mechanisms meant to keep
public services provision in check. Other legislation comprises bills that do
not fit in the previous categories, such as changes in street names, motions to
honor citizens or groups, and legislative and internal city council procedures.
We provide further information about the legislation dataset and the city
council survey in sections A.13 and A.14 of the Supplementary Materials.

The last set of variables quantifies social welfare. Our focus is on the impact
of legislature size on education and health care, which are the two most
important public services the municipal government should provide to citizens.
According to Brazilian Constitution, municipalities must spend 25\% of their
budget on education and 15\% on health care. Consequently, the Ministry of
Health and the Ministry of Education provide readily available, fine-grained
data on health care and education at the municipal level, which we leverage in
this study. We measure access to education with the average enrollment in
primary (K--4) municipal public schools. We assess education quality with the
Education Development Index (Índice de Desenvolvimento da Educação
Básica--IDEB), which the Ministry of Education has issued biannually since
2007. The index is a weighted average of student scores in Portuguese and
Mathematics (0--10), multiplied by the harmonic mean of public school
promotion rates in a given year (0--100). To test the effect of the 2004
ruling on health care, we collect data on infant mortality per thousand
children born and on postneonatal infant mortality, which is the death of
infants who survived more than 28 days, but died before reaching one year
of age. The source of all health care variables is the Brazilian Ministry
of Health Data Center (DataSUS).

Together, this evidence provides a clear picture of local-level politics in
Brazil, from the political mechanism to the municipal welfare consequences of
the council size ruling. Summary statistics and variable descriptions follow in
section A.2 of the Supplementary Materials.

# Empirical Strategy
\label{sec:empstrat}

We employ a regression discontinuity design (RDD) using the population
thresholds established by the 2004 Supreme Electoral Court resolution as
treatment indicators. Figure \ref{fig:runningvarplot} displays the distribution
of municipalities within each city council size. These population thresholds
remained in effect until 2009 when Congress amended the Brazilian Constitution.
This study covers 2005 to 2008, which constitutes the full mayoral and
city-council term following the 2004 Brazilian elections.

Our identification strategy relies on three assumptions. First, no municipality
should be able to self-select into each side of the discontinuity. Second, as
we estimate a sharp RDD, the council size should increase precisely as the law
mandates. Third, this design assumes that the pre-treatment variables,
collected before the 2003 Supreme Federal Court decision, were not affected by
the 2004 ruling on council size. Moreover, for all the actual models estimated
here, we fitted placebo cutoffs to check whether the estimation technique leads
to consistent results. The placebo cutoffs are the dashed lines in Figure
\ref{fig:runningvarplot}.

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=6in,angle=0]{../figures/distrSeats.pdf}}
      \caption{Distribution of Municipalities by Population and City Council Size}
      \bigskip
      \label{fig:runningvarplot}
    \end{minipage}
  \end{center}
\end{figure}

For the first assumption, in section A.3 of the online supplemental materials,
we assess distribution imbalances running the \citet{McCrary2008} and
\citet{cattaneo2019simple} tests. The tests show no distribution imbalances.

For the second assumption, running RDD in multiple thresholds may lead to
inconsistent estimates when there are differences in the distribution of cases
around the cutoffs \citep{cattaneo2016interpreting, eggers2018regression,
bertanha2020regression}. In the dataset, there are twelve discontinuities. At
each threshold, there is a sharp increase of one councilor, from 9 to 21. When
running an RDD on these discontinuities, a consistent estimator should fit the
exact change of one councilor. However, if we pool all the discontinuities
together with no correction, we find an increase in council size of 1.63
councilors (see the first estimate in Table \ref{validitycheck}). This
imbalance occurs because when we combine all the discontinuities, the model
implies that municipalities below the 47,619 cutoffs (9 to 10 councilors) are
comparable to changes in towns right above the 571,428 cutoffs (20 to 21
council members). That is incorrect because the comparison is inconsistent in
council size, and these municipalities diverge in many other aspects.

To avoid this problem, in section A.5 of the Supplemental Materials, we show by
simulation that adding controls, especially the variables responsible for the
multiple threshold assignments (in our case, the population in 2003), improves
the consistency and efficiency of the estimates. Relying on this fact, we add
five controls to the estimates: population in 2003; GDP per capita; the number
of seats in 2000; year; and a dummy for Northeast municipalities. The
reason for the first variable is the multiple threshold assignments. We add GDP
per capita because wealthier cities tend to have better public services. The
number of seats in 2000 intends to control that some municipalities could
experience a change provoked by the previous council size, confounding our
estimates. The year and the Northeast dummies were added to improve efficiency.
Adding controls improves efficiency in RDD and also enhances the consistency in
the multiple thresholds RDD \citep{calonico2019regression}.

For the last assumption, there should be no changes in pre-treatment
covariates. These pre-treatment covariates are measured before the primary
outcomes. As the threshold rule selected by the Brazilian Electoral Court was
unpredictable, it should not detect any variation before the thresholds were in
place.

Table \ref{validitycheck} displays the results of the validity checks. We
present the first-stage regressions at the top of the Table, with and without
covariates, for both real and placebo cutoffs. At the bottom, we estimate the
pre-treatment covariate balance. Furthermore, in the Supplementary Materials,
we run placebo checks for all the models using placebo cutoffs within the real
ones. The results are available in sections A.6 to A.12 of the Supplementary
Materials.

\singlespacing

\begin{table}[H] \centering 
  \caption{Research Design Validity Check} 
  \label{validitycheck} 
    \scalebox{0.9}{
  \begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{4}{c}{\textbf{Panel A: Validity Check --- Treatment Effect on City Council Size}} \\ \cline{2-5}
   & \multicolumn{1}{c}{\small{(1)}} & 
 \multicolumn{1}{c}{\small{(2)}} & 
 \multicolumn{1}{c}{\small{(3)}} & 
 \multicolumn{1}{c}{\small{(4)}} \\

 & \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} & 
 \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} \\ 
  & \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} & 
  \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} \\
  & \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
\\[-1.8ex] 
 LATE & 1.63$^{***}$ & -0.31$^{**}$ & 1.00$^{***}$ & -0.00 \\ 
  & (0.51) & (0.14) & (0.0004) & (0.0000) \\ 
\hdashline \\[-1.8ex]   
 N Left & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} \\ 
Eff N Left & \multicolumn{1}{c}{199} & \multicolumn{1}{c}{638} & \multicolumn{1}{c}{49} & \multicolumn{1}{c}{886} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{385} & \multicolumn{1}{c}{51} & \multicolumn{1}{c}{477} \\ 
BW Loc Poly & 8.717 & 6.700 & 3.008 & 8.629 \\ 
BW Bias & 13.620 & 12.114 & 5.008 & 13.466 \\ 
\midrule 
 & \multicolumn{4}{c}{\textbf{Panel B: Validity Check --- Pre-Treatment Variables}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(5)}} & 
 \multicolumn{1}{c}{\small{(6)}} & 
 \multicolumn{1}{c}{\small{(7)}} & 
 \multicolumn{1}{c}{\small{(8)}} \\
  & \multicolumn{1}{c}{\small{Number of}} & 
 \multicolumn{1}{c}{\small{Population}} & 
 \multicolumn{1}{c}{\small{GDP}} & 
 \multicolumn{1}{c}{\small{\% of Poverty}} \\ 
  & \multicolumn{1}{c}{\small{Seats 2000}} & 
  \multicolumn{1}{c}{\small{2000 Census}} & 
  \multicolumn{1}{c}{\small{2000 Census}} &
   \multicolumn{1}{c}{\small{2000 Census}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
 LATE & 0.17 & -2057.60 & 0.06 & -9.72 \\ 
  & (0.67) & (1926.33) & (0.18) & (6.27) \\ 
  \hdashline \\[-1.8ex]
 N Left & \multicolumn{1}{c}{5178} & \multicolumn{1}{c}{5131} & \multicolumn{1}{c}{5131} & \multicolumn{1}{c}{5131} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} \\ 
Eff N Left & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{219} & \multicolumn{1}{c}{200} & \multicolumn{1}{c}{194} \\ 
Eff N Right & \multicolumn{1}{c}{147} & \multicolumn{1}{c}{157} & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{142} \\ 
BW Loc Poly & 8.970 & 9.368 & 8.739 & 8.530 \\ 
BW Bias & 13.860 & 15.029 & 14.752 & 13.035 \\ 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $ ^{***} $ p $ < $ .01; $ ^{**} $ p $ < $ .05; $ ^{*} $ p $ < $ .1. RD local linear estimates using \citet{calonico2019regression} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population, GDP per capita, number of seats in 2000, year, and dummy for Northeast region. \textit{N Left} and \textit{N Right} represent the total number of observation in the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

\doublespacing

Note that without the covariates, the first stage overestimates the change in
the primary treatment and underestimates the effect in the placebo regressions.
When we add the covariates, the results improve and the point estimates are now
one for the primary model and zero for the placebo regressions. These are the
correct values. Moreover, Panel B of Table \ref{validitycheck} shows that the
pre-treatment covariates remain unchanged around the thresholds. These tests
reinforce the validity of our research design.

# Empirical Results
\label{sec:results}

To test the predictions from our formal model, we present four sets of
empirical results. First, we confirm the bargaining costs hypothesis and find
that the additional legislator also promotes more appointments of bureaucrats.
This aligns with the councilors' descriptions of the strategies mayors employ
to consolidate support from city councilors. Second, we show that
municipalities with an extra city councilor approve more legislation about
public goods provision in general and about education and health care in
particular. Third, we show that all these pieces of evidence result in
improvements in infant mortality and elementary school enrollment, two vital
services for welfare. Finally, we also show that our results are robust to
alternative channels for the influence of legislature size on welfare,
demonstrating that representation of women and non-whites, competitiveness, and
legislative productivity fail to receive empirical support.

## Council Size, Bargaining Costs, and Job Appointments
\label{sub:bargcosts}

The main premise of our model is that larger legislature size produces lower
bargaining costs when the chances of electing government-aligned politicians
are sufficiently high. Moreover, the main bargaining chip between the executive
and the legislative is job appointments, even more so when the city councilors
are aligned with the mayor. As job appointments are an important factor for
local governance, we analyze the number of politically appointed bureaucrats,
career bureaucrats, and councilors' assistants. A higher number of political
appointees is associated with better service provision in Brazil, as their
tenure in office depends directly on the survival of the politician supporting
them \citep{toral2019benefits}. Career bureaucrats pass through a rigorous
examination, limiting their usefulness as bargaining chips between city
councilors and mayors. Moreover, career bureaucrats cannot be fired once hired,
which considerably constrains politicians' pressure on them. Finally, city
councilors who want to influence policy may employ more staff members and
assistants to their offices. This could influence policymaking through city
councilors putting pressure on the bureaucracy, defying our hypothesized
mechanism. Our theoretical expectations for these regressions are that larger
council size affects the political alignment of city councilors and the number
of politically appointed bureaucrats. It should not affect the number of career
bureaucrats or city council assistants, as it is more valuable to have a
political appointee that politicians can pressure, working directly with
voters. Table \ref{tab:bargain} tests the first hypothesis about the nexus
between city council size and bargaining costs and explores the consequences of
larger legislatures on the municipal bureaucracy.

\begin{table}[!htpb] \centering 
  \caption{Political Effects of Increasing the Number of Legislators} 
  \label{tab:bargain} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} &
 \multicolumn{1}{c}{\small{(3)}} & 
 \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Mayoral}} & 
 \multicolumn{1}{c}{\small{Num. Politically}} & 
 \multicolumn{1}{c}{\small{Num. Career}} & 
 \multicolumn{1}{c}{\small{Num. Councilor}} \\
 & \multicolumn{1}{c}{\small{Coalition Size}} & 
 \multicolumn{1}{c}{\small{Appointed Bureaucrats}} & 
 \multicolumn{1}{c}{\small{Bureaucrats}} & 
 \multicolumn{1}{c}{\small{Assistants}} \\[1ex]  \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
 LATE & 0.91$^{*}$ & 104.49$^{*}$ & 71.57 & 2.09 \\ 
  & (0.50) & (61.53) & (218.12) & (4.37) \\ 
    \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5179} & \multicolumn{1}{c}{15536} & \multicolumn{1}{c}{15531} & \multicolumn{1}{c}{5179} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{1027} & \multicolumn{1}{c}{1027} & \multicolumn{1}{c}{344} \\ 
Eff N Left & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{372} & \multicolumn{1}{c}{513} & \multicolumn{1}{c}{99} \\ 
Eff N Right & \multicolumn{1}{c}{162} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{388} & \multicolumn{1}{c}{101} \\ 
BW Loc Poly & 10.025 & 6.242 & 7.695 & 5.134 \\ 
BW Bias & 15.889 & 10.243 & 11.780 & 8.902 \\ 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

Column (1) shows a 91\% chance of the extra legislator belonging to the mayoral
coalition. The model hypothesis stated that the probability of electing a city
councilor aligned with the mayor is sufficiently high. However, sufficiently
high can be ambiguous, and it depends on the intensity of the political
disagreement and the current legislature size. In any case, 91\% chance is a
considerably large quantity, and the extra city councilor is going to the
mayoral coalition nine in every ten legislature size increases. This
strengthens the political position of the mayor and diminishes bargaining
costs.\footnote{The effect of city council size on the number of mayor-aligned
politicians stems from the rounding rule adopted by the proportional
representation system in Brazil. The rounding rule favors the coalitions with
more votes. Most of the time, the coalition of the elected mayor is the one
that receives more votes.} Column (2) demonstrates that adding one councilor
increases appointed employees by around 104 extra employees per additional city
councilor. When negotiating with city councilors, these employees are a great
bargaining chip, and they are also associated with better service provision
\citep{toral2019benefits}. As expected, legislature size has a null effect on
the number of career bureaucrats (Column 3) and councilors' staff members
(Column 4).

## Approved Legislation and City Councilor's Survey
\label{sub:survey}

We built two novel datasets to provide evidence for our mechanism. First, we
ran a survey with former city councilors, asking how mayors consolidate their
coalition support within the city council.\footnote{We provide further
information about the city council survey in section A.14 of the Supplementary
Materials.} Second, we collected 346,553 laws approved in 63 municipalities
within ten thousand inhabitants away from the city council size thresholds. We
hand-coded 3,466 laws into four categories: public goods, oversight, education
and health care-related legislation, and other legislative duties. After
hand-coding these laws, we used a Supporting Vector Machine (SVM) algorithm to
classify the remainder laws. From all of the algorithms we tested, SVM achieved
the highest performance.

\begin{figure}[htb!]
  \begin{center}
  \caption{Mayoral Instruments for Securing Councilors' Support}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=6in,angle=0]{../figures/strategySupportMayors.pdf}}
      \bigskip
      \label{survelec}
    \end{minipage}
  \end{center}
\end{figure}

We assume that city councilors want to extract rents and to get reelected.
The rents come from continuing in the repeated game, and the partisanship
discounts that benefit. Figure \ref{survelec} shows that 65.9 percent of
councilors mentioned that mayors use job appointments to consolidate their
coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed
that mayors use personal favors, such as shifting resources to places where the
councilors have electoral support, to solidify their governing coalition.
Granting favors works as rent transfers for councilors. Job appointments are
also convenient for the mayors because patronage in Brazil has positive results
in terms of welfare \citep{toral2019benefits}.

\begin{figure}[H]
  \begin{center}
  \caption{Legislation Approval Patterns in the Municipalities Closer to the City Council Size Thresholds}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=6in,angle=0,height=6.5in]{../figures/lawsApproved.pdf}}
      \bigskip
      \label{LEGAPPR}
    \end{minipage}
  \end{center}
\end{figure}

In Figure \ref{LEGAPPR}, we show that the per capita number of laws related to
local public goods provision increases in towns that receive the extra
legislator, vis-à-vis baseline cities that did not get an additional councilor.
One additional city councilor increased the per capita amount of laws on public
goods by 15.0 percent. Also, the number of laws that focus on health care and
education increased by 19.7 percent. Looking specifically into health care and
education legislation, the proportion of laws discussing public goods increased
by 15.7 percent, showing that policy proposals by both the mayor and the
councilors shift toward local and citywide public goods.\footnote{Public goods
here are not restricted to citywide goods. Any service that benefits citizens,
regardless of whether they are targeted to citizens or not, is considered
public goods. For example, building a health clinic in a particular district
can virtually exclude the citizens that are not living in the neighborhood.
However, we are still considering these as public goods.}

## The Effect of Legislature Size on Welfare
\label{sub:effects}

We now study the impact of legislature size on education and health care. Table
\ref{tab:outcomes} shows that increasing the city council by one legislator
generates welfare improvement in municipalities. Regarding health care,
increasing the legislature size decreases infant mortality in 2.01 deaths per
1,000 children born. This impact is significant in absolute terms and
represents an effect of 0.14 in standard deviations. Moreover, postneonatal
mortality decreases by 0.90 casualties per 1,000 births, which remains around
the 0.14 standard deviations improvement as infant mortality. Infant mortality
is a primary outcome for welfare, and it correlates with several other
indicators of well-being \citep{world2005health}. The fact that infant
mortality improved in the Brazilian case shows that the increased mayor's
support mechanism results in large welfare effects. As expected, placebo
regressions were statistically insignificant.

In terms of education, with the addition of a city councilor, enrollment in
elementary schools increased by 2.58 children on average per school. This
represents a change of around 0.20 standard deviations. Besides the influence
of the mayoral support, enrollment can also directly impact the councilor's
representation: Councilors can pressure the bureaucracy to accommodate a few
extra students. With support from the mayor, they can help voters get access to
schools, improving the welfare of voters. Education quality remains unchanged,
but with more students in the classroom, it could potentially decrease. As
expected, the placebo regressions are insignificant.

\begin{table}[!htpb] \centering 
  \caption{Legislature Size and Welfare} 
  \label{tab:outcomes} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Infant}} & 
 \multicolumn{1}{c}{\small{Postneonatal}} &
  \multicolumn{1}{c}{\small{School}} & 
  \multicolumn{1}{c}{\small{Education}} \\ 
   & \multicolumn{1}{c}{\small{Mortality}} & 
 \multicolumn{1}{c}{\small{Mortality}} &
  \multicolumn{1}{c}{\small{Enrollment}} & 
  \multicolumn{1}{c}{\small{Quality}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & -2.01$^{**}$ & -0.90$^{*}$ & 2.52$^{***}$ & -0.03 \\ 
  & (0.79) & (0.48) & (0.80) & (0.13) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{12299} & \multicolumn{1}{c}{5441} & \multicolumn{1}{c}{10156} & \multicolumn{1}{c}{8597} \\ 
N Right & \multicolumn{1}{c}{1029} & \multicolumn{1}{c}{672} & \multicolumn{1}{c}{686} & \multicolumn{1}{c}{669} \\ 
Eff N Left & \multicolumn{1}{c}{501} & \multicolumn{1}{c}{324} & \multicolumn{1}{c}{206} & \multicolumn{1}{c}{409} \\ 
Eff N Right & \multicolumn{1}{c}{381} & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{291} \\ 
BW Loc Poly & 7.527 & 7.564 & 5.425 & 9.05 \\ 
BW Bias & 11.95 & 11.01 & 10.58 & 15.107 \\ [1.0ex] 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

## Robustness to Alternative Mechanisms

In this paper, we show that the provision of education and health care in
Brazil improves with larger legislatures. This is because the chances of
electing a city councilor aligned with the mayor are 91 percent, thanks to the
rule of distributing seats adopted in the country. In the Supplemental
Materials, we also run regressions in placebo cutoffs between the actual
cutoffs, showing that our estimates are consistently null when we are indeed
supposed to find insignificant results.

However, other pathways may also explain the nexus between city council size
and welfare. First, improvements in representation could change the provision
of health care and education. Increasing legislature size without changing the
population makes councilors represent fewer people. For example, in a town with
50 thousand voters and ten councilors, each councilor represents five thousand
voters. If the same city has eleven councilors, then each councilor represents
4.5 thousand voters. This facilitates the representation of minorities, as they
have to gather fewer votes to elect a city councilor. A minority representative
can advocate for policies that would improve the welfare of marginalized groups
\citep{ChattopadhyayDuflu2004, chin2011redistributive, pande2003can}. Diversity
could also improve governance in municipalities \citep{rugh2011provision}. For
instance, more women in public offices may shift service provision toward
welfare expenditure \citep{brollo2016happens, duflo2012women,
funk2019representative, hernandez2018women}. Therefore, an alternative
hypothesis to explain the increased service provision is that larger
legislatures improve female and minority representation in Brazil.\footnote{For
a description of the barriers faced by non-white politicians in Brazil, see
\citet{janusz2021race}.}

Second, more seats could increase the competitiveness in the election, making
politicians provide more services \citep{besley2010political}. In a political
system with around thirty parties and a sample average of almost ten parties in
the municipalities we study, candidates per seat can increase exponentially.
According to the Brazilian electoral law, each party can sponsor 1.5 candidates
per seat. If a party forms a pre-electoral coalition, this number increases to
two times the number of seats \citep{coneglian2008lei}. An election in a small
town, with ten councilors, may end up with 600 candidates: (30 parties) x (two
times the number of seats) x (ten seats). Therefore, an alternative explanation
for better service provision is that a larger council size increases the
competitiveness of the election.\footnote{Despite the common assumption that
competitiveness improve welfare, \citet{boulding2014political} and
\citet{gottlieb2019countervailing} show that more competition can increase
budgetary pressure, lower social policy, lead to bargaining inefficiencies.}

Finally, the mayor and the legislature carry out the provision of services.
Similar to what happens in firms, an extra employee can increase the firm
productivity frontier by improving cooperation and synergistic outputs
\citep{alchian1972production, holmstrom1982moral}. A similar argument would
posit that the increase in council size increases the capability for
municipalities to provide public services. As legislative productivity
manifests itself as legislation, an alternative hypothesis to explain the
increased service provision is that more legislation is being approved by the
city council.

\begin{table}[!htbp] \centering 
  \caption{Robustness: Effects of Legislature Size Changes on Representation, Competitiveness, and Legislative Productivity} 
  \label{tab:robust} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Num. Female}} & 
 \multicolumn{1}{c}{\small{Num. Non-white}} &
  \multicolumn{1}{c}{\small{Candidates}} & 
  \multicolumn{1}{c}{\small{Prop. Laws}} \\ 
   & \multicolumn{1}{c}{\small{councillors}} & 
 \multicolumn{1}{c}{\small{councillors}} &
  \multicolumn{1}{c}{\small{Per Seat}} & 
  \multicolumn{1}{c}{\small{Approved Council}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & 0.22 & 0.60 & -0.19 & -0.02 \\ 
  & (0.38) & (0.85) & (1.11) & (0.08) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5183} & \multicolumn{1}{c}{239} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{3424} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{158} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{269} \\ 
Eff N Left & \multicolumn{1}{c}{195} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{179} & \multicolumn{1}{c}{227} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{132} & \multicolumn{1}{c}{141} \\ 
BW Loc Poly & 8.623 & 2.958 & 8.144 & 10.976 \\ 
BW Bias & 13.565 & 4.922 & 12.018 & 16.708 \\[1.0ex] 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

Table \ref{tab:robust} tests whether our results could stem from these
different mechanisms. Columns (1) and (2) test the hypotheses that better
representation of under-served groups could improve welfare. As we can see,
larger city councils do not increase the representation of females or
non-whites. Column (3) shows that the competition per seat in municipalities
with larger city councils remain statistically equal to zero. Finally, Column
(4) shows that the proportion of legislation passed by the city council remains
unchanged when council size increases.

# Discussion

This paper shows that partisanship considerably decreases bargaining costs and
improves citizen welfare. The crucial actors in providing public goods are the
mayors, but they cannot do it alone. Legislatures are a solid check to the
mayor's policy. Bargaining with them in a country plagued by corruption can be
costly \citep{bertholini2017price}. Therefore, an extra ally in the city
council may help ease the bargaining and increase social welfare.

Increasing the council size results in gains in elementary school enrollment
and in lower infant mortality. Education quality remains unchanged, but
increasing enrollment without affecting quality should be interpreted as a
positive outcome. The analysis of the approved legislation shows that the
frequency of laws on local public goods increases in municipalities above the
council size thresholds. Finally, our survey data demonstrate that councilors
and mayors agree about appointing more public servants to the public
administration.

\citet{Ames2001} argues that the Brazilian democracy is prone to deadlocks
because the executive has its hands tied by legislators that demand consistent
provision of particularistic and private goods. Indeed, the most successful
presidents after 1995 had to build multiparty coalitions by negotiating with
parties \citep{limongi1998bases, raile2011executive}. Despite evidence that
side payments cemented these coalitions, many social and economic policies have
since been implemented. For instance, education and health care notably
improved during the Cardoso and Lula governments \citep{acemoglu2012nations}.
Presidents who challenged the coalition system did not succeed. For instance,
former President Rousseff was impeached in her second term. Bolsonaro can be
considered one of the most ineffective presidents in Brazilian history.

Municipalities work similarly in terms of the incentives for
legislative-executive bargaining. The stakes are lower, and the negotiations
involve only one legislative branch. Still, in a system such as Brazil's, with
more than thirty-five registered parties, mayors face substantial challenges to
the support for their agendas. Having extra help in the council can tilt the
bargaining towards cheaper governing costs, thus improving public goods
provision.

However, we do not argue that these benefits result from the absence of
corruption or clientelism. Both the formal model and empirical evidence
provided by \citet{britto2020corruption} show that corruption increases with
legislature size. As the model shows, city councilors receive a side transfer
from mayors to support their agenda. Mayors also divide the resources saved in
the bargain into public goods and rents. The rate for the division depends on
the marginal changes of the rents and public goods, but these two components
receive the extra resources. In this case, the mayor's rents may be primarily
corruption, and the councilor's transfers can be clientelistic, aimed at
improving the councilor's reelection chances.

Our findings have implications that extend well beyond Brazil. A general form
of our argument is that it is easy for a politician to negotiate with members
of her in-group \citep{abbink2019, g9030057}. Specifically, several countries
have separate executive and legislative powers, such as the federal level of
presidential systems, state governments, and local city councils. As these
political bodies are influenced by political allegiances, the dynamics we
expose here may be generalizable. Depending on legislative support, changing
the legislature size could increase the chance of electing politicians that are
not aligned with the executive, hence decreasing welfare. Despite the
particular effects on political support and welfare, the logic described in
this paper would remain relevant for empirically accessing the effect of
enlarged legislatures on welfare. Our findings may also help scholars
understand why other countries in Latin America and Asia can generate effective
governance despite having strong executives and large coalitions
\citep{pereira2012surprising}.

This paper opens several questions for future analysis. First, if the mechanism
we suggest is indeed at work, one may ask what should be the city council's
checking prerogatives, as they significantly reduce the provision of public
services. The evidence so far is inconclusive, and \citet{poulsen2018political}
show that checks for the mayors may also improve service provision. Second,
future research could evaluate whether additional legislators help mayors gain
access to federal or state transfers in countries where resources are centrally
provided, such as Japan. As suggested by \citet{catalinac2020tournament}, in
these situations the national government can create a tournament between
municipalities and provide resources to those which offer more electoral
support. Scholars might want to analyze whether this tournament structure
interacts with the mechanisms we suggest in this paper. Third, scholars might
want to investigate whether our findings remain valid under different
conditions. It is possible that the mechanisms we present here are non-linear,
with additional legislators having a positive or negative impact on welfare
depending on certain electoral characteristics. Finally, it remains uncertain
how other legislative features affect citizen well-being. For example, city
councils may vary in size, monetary compensation for councilors, committee
structure, and internal power structure. Understanding how legislature size
interacts with other features would improve our knowledge about how local
legislatures impact public service delivery. These questions are crucial for
institution design and welfare in developing democracies.

\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}