### 2022-05-16 Memo/memo.Rmd
update memo and appendix

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: Response to Reviewers
date: \today
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: single
papersize: a4paper
bibliography: ../article/references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: no
    keep_tex: no
    toc: no
    toc_depth: 3
    template: memo-template.latex
    md_extensions: +raw_attribute
---

\noindent Dear Editors and Reviewers,
\vspace{.5cm}

\noindent We would like to thank you all for the opportunity to revise our
manuscript, "Legislature Size and Welfare: Evidence from Brazil" (Ms. No.
AJPS-44979). We have made several revisions to the paper along the lines
suggested by the Editors and Reviewers and we believe the manuscript has
improved significantly as a result.

Following Reviewers 1 and 2, we have incorporated additional literature into
the manuscript, which has given the paper a clearer theoretical framework
through which our results can be interpreted. We have also discussed the scope
conditions of our findings in further detail and included a new paragraph
explaining why newly-elected councilors tend to join pro-mayor coalitions, thus
addressing two important concerns raised by Reviewer 3. We believe these
changes have made the paper clearer and more comprehensive.

In response to comments from Reviewer 1, we have revised our Methods section to
clarify our selection of independent variables, including the use of covariates
to improve the efficiency and consistency of our estimates. Additionally, we
have included a new section in the Supplemental Materials to explain how we
conducted our survey and how it contributes to our results. We believe these
additions also give the paper a better theoretical frame to help readers sort
through the results we present.

Below we discuss in more detail these changes and the several others we have
made to the manuscript in response to comments from the Editors and Reviewers.
We thank the Editors again for the opportunity to revise the manuscript, and
the Reviewers for their extremely helpful comments.

\vspace{.5cm}
\noindent Sincerely,

\vspace{.2cm}
\noindent The Authors

# Reviewer 1 Comments and Responses

\noindent 1) Reviewer 1 writes: "Overall the literature review is strong. I
would recommend a couple of additions. Include among them is an important work
the authors should cite on legislative size and spending and that is Gilligan
and Matsusaka's 2001 article entitled 'Fiscal Policy, Legislature Size, and
Political Parties: Evidence from the First Half of the Twentieth Century.'
Their findings are relevant for this study and should be referenced. In
addition, the recently published book \textit{Population and Politics: The
Impact of Scale} by Gerring and Veenendaal which has a comprehensive analysis
of this topic should also be incorporated into the literature review."

\vspace{.2cm}

\noindent \textbf{Response:}

\begin{quote}

We thank Reviewer 1 for these two relevant suggestions. We now cite Gilligan and
Matsusaka's 2001 article three times in our paper. We mention it twice in our
Introduction, first on page 2: ``\textit{On the one hand, enlarged legislatures
may reduce political particularism and increase overall government spending,
which can lead to more public goods  \citep{gerring2020population,
GilliganMatsusaka2001, Weingastetal1981}}'', and then in the end of page 4:
``\textit{The corollary of the "law of $1/n$" is that increasing legislature
size promotes excessive public spending, regardless of the partisan makeup of
the legislature \citep{GilliganMatsusaka2001}}''. We also cite it in our Theory
section (page 7): ``\textit{Yet, \citet{GilliganMatsusaka2001} find that in the
US legislature size is associated primarily with higher spending on
geographically concentrated benefits, such as education and highways}''.

We cite Gerring and Veenendaal's book four times. We mention it twice on page 2:
``\textit{Legislature size is known to have a significant effect on collective
action strategies, the bargaining power between state branches, government
spending, and representation \citep{crain1979cost, freire2021legislature,
gerring2020population, PrimoSnyder2008, rogers2002free, Weingastetal1981}.
However, how legislature size impacts social welfare remains less understood. On
the one hand, enlarged legislatures may reduce political particularism and
increase overall government spending, which can lead to more public goods
\citep{gerring2020population, GilliganMatsusaka2001, Weingastetal1981}}''. We
mention the book again on page 4: ``\textit{Following the seminal contribution
by \citet{Weingastetal1981}, the "law of $1/n$" posits that larger legislatures
increase the legislative pressure for providing targeted public goods to local
supporters at the expense of the entire polity \citep{freire2021legislature}. This
question has become more relevant since legislature size has increased
consistently over time \citep{gerring2020population}}''. Finally, we also cite
it on page 6: ``\textit{Larger legislatures improve representation as both the
election thresholds and the ratio of population-per-politician decrease, which
facilitate the electoral success of ethnic and political minorities
\citep{allenrepresentation, gerring2020population}}.''

\end{quote}

\noindent 2) Reviewer 1 writes: "In examining the relationship between
legislative size and welfare why were education and health care selected as the
independent variables? Was that decision based on data availability or were
there another basis for the decision? It is not clear in the text of the
manuscript."

\vspace{.2cm}

\noindent \textbf{Response:}

\begin{quote}

We thank Reviewer 1 for the opportunity to clarify the selection of these two
sets of variables for our paper. We included an explanation for our choice on
page 17: "\textit{Our focus is on the impact of legislature size on education
and health care, which are the two most important public services the municipal
government should provide to citizens. According to the Brazilian Constitution,
municipalities must spend 25\% of their budget on education and 15\% on health
care. Consequently, the Ministry of Health and the Ministry of Education provide
readily available, fine-grained data on health care and education at the
municipal level, which we leverage in this study}."

\end{quote}

\noindent 3) Reviewer 1 writes: "The details of the online survey conducted of
Brazilian city councilors are lacking. What is the empirical contribution to
the study? Beyond the online format what were the other administrative details
involved with collecting that data?"

\vspace{.2cm}

\noindent \textbf{Response:}

\begin{quote}

We agree with Reviewer 1 that we needed to provide more information on the
online survey. Therefore, we have detailed our procedures in Section A.14 of the
Supplementary Materials (pages 22-23), and we also reference the survey on pages
17 and 22 (footnote 13) in the main article. We rewrote paragraph 2 of Section
6.2 to clarify the importance of the survey results in our paper. It reads as
follows: ``\textit{We assume that city councilors want to extract rents and to
get reelected. The rents come from continuing in the repeated game, and the
partisanship discounts that benefit. Figure 2 shows that 65.9 percent of
councilors mentioned that mayors use job appointments to consolidate their
coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed
that mayors use personal favors, such as shifting resources to places where the
councilors have electoral support, to solidify their governing coalition. This
provides further evidence that mayors do offer city councilors the benefits they
demand. In this regard, granting favors works as rent transfers for councilors.
Job appointments are also convenient for the mayors because patronage in Brazil
has positive results in terms of welfare \citep{toral2019benefits}.}''

\end{quote}

\noindent 4) Reviewer 1 writes: "The models include a dummy variable for the
northeast region. What specifically about this region would require adding it
to the model. I am generally skeptical about including such a variable
particularly since clustered standard errors are being estimated for municipal
level. Is there logical reason to control for regional effects here? Very
curious modeling decision by the authors."

\vspace{.2cm}

\noindent \textbf{Response:}

\begin{quote}

We thank Reviewer 1 for the opportunity to clarify the selection of control
variables in our empirical analysis. The inclusion of controls in our model has
two purposes. The first one is to improve the consistency of the estimates. In
Section A.5 of the Appendix, we use simulation to show that the inclusion of
covariates enhances the consistency of our point estimates. The simulation
indicates improved consistency for several distributions of the running
variables and different treatment effects. Section 5 of the manuscript
demonstrates that it improves consistency in both the main estimates and the
placebo ones. Table 1 (page 20, reproduced below) shows that without controls,
we would overestimate the treatment effects (finding an increase of 1.63
legislators instead of an increase of exactly one legislator), and underestimate
the placebo treatment effects (finding a significant decrease of 0.31
legislators where there was supposed to be no effects in legislature size). This
substantial improvement is because having control variables makes the
subpopulations in each threshold more comparable across all the thresholds,
avoiding comparing cases to the left (right) of one given threshold with points
in the right (left) of a different threshold.

\begin{table}[H] \centering 
  \caption{Research Design Validity Check (reproduction from the main paper)}
    \scalebox{0.9}{
  \begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{4}{c}{\textbf{Panel A: Validity Check --- Treatment Effect on City Council Size}} \\ \cline{2-5}
   & \multicolumn{1}{c}{\small{(1)}} & 
 \multicolumn{1}{c}{\small{(2)}} & 
 \multicolumn{1}{c}{\small{(3)}} & 
 \multicolumn{1}{c}{\small{(4)}} \\

 & \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} & 
 \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} \\ 
  & \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} & 
  \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} \\
  & \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
\\[-1.8ex] 
 LATE & 1.63$^{***}$ & -0.31$^{**}$ & 1.00$^{***}$ & -0.00 \\ 
  & (0.51) & (0.14) & (0.0004) & (0.0000) \\ 
\hdashline \\[-1.8ex]   
 N Left & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} \\ 
Eff N Left & \multicolumn{1}{c}{199} & \multicolumn{1}{c}{638} & \multicolumn{1}{c}{49} & \multicolumn{1}{c}{886} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{385} & \multicolumn{1}{c}{51} & \multicolumn{1}{c}{477} \\ 
BW Loc Poly & 8.717 & 6.700 & 3.008 & 8.629 \\ 
BW Bias & 13.620 & 12.114 & 5.008 & 13.466 \\ 
\bottomrule
\end{tabularx}
\end{threeparttable}}
\end{table}

The second reason is that adding control variables improves the efficiency of
our estimates \citep{calonico2019regression}. We add five controls to the main
specifications. We include the 2003 population because it conditions the
multiple threshold assignments. We add GDP per capita to control for Brazil's
economic expansion during the 2000s. We also add the number of seats in 2000 to
ensure that our results come from the current legislature size and not from past
seat allocations. Year was included in the models to control for seasonal
effects. Finally, we add a dummy for Northeast states. The Northeast region
received considerable attention from the Federal Government in the 2000s.
According to \citet{de2015politica}, in 2003 the Lula administration recreated
the Superintendency for the Development of the Northeast (SUDENE), investing BRL
84 Billion from 2003 to 2010. This amounts to around 70% of all targeted
regional investments. One example of these investments is that the Federal
Government created about 100 new universities in the Northeast region. Moreover,
infant mortality in the region decreased from 22.4 deaths per 1000 infants born
in 2005 to 19.6 deaths per 1000 infants born in 2008 due to larger federal
investments. In turn, education quality improved from 2.69 in 2005 to a 3.14
score in 2007 (0 to 10 quality of education score). For this reason, adding a
dummy for Northeast states is essential for ensuring the consistency and the
efficiency of our estimates.

To reflect these pieces of information, we rewrote paragraph five in Section 5
of the manuscript to (page 19):

``[...] \textit{we add five controls to the estimates: population in 2003; GDP
per capita; the number of seats in 2000; year; and a dummy for Northeast
municipalities. The reason for the first variable is the multiple threshold
assignments. We add GDP per capita because wealthier cities tend to have better
public services. The number of seats in 2000 intends to control that some
municipalities could experience a change provoked by the previous council size,
confounding our estimates. The Northeast dummy was added to control for the fact
that the Lula administration heavily invested in the region from 2003 until
2010, improving several development indicators \citep{de2015politica}. The years
were added to improve efficiency and control for seasonal effects. Moreover,
adding controls improves efficiency in RDD \citep{calonico2019regression} and
also enhances the consistency in the multiple thresholds RDD.}''

And we added the discussions here to the Appendix at the end of Section A.5.

We want to highlight further that this decision to add controls in our case is
also because of consistency. Even though clustered standard errors may provide
the best standard errors for our model, this would not affect the consistency.
And the consistency here relies on making the subpopulations comparable, i.e.,
the subpopulation of towns between 9 and 10 legislators as comparable as the
subpopulation of towns between 10 and 11 legislators, and so on.

\vspace{.2cm}

We would like to thank Reviewer 1 for their helpful comments and suggestions.
\end{quote}

# Reviewer 2 Comments and Responses

\noindent 1) Reviewer 2 writes: "One minor issue for the author(s) to consider
in future research: on p. 14 the paper states, "On average, municipal taxes
account for only 5\% of the local budgets, while mandatory transfers from state
and federal governments correspond to the remaining 95\%. This reduces the
ability of municipal legislators to effectively increase local spending, which
makes improvements in local services dependent only on the composition of their
local investments." There is an alternative mechanism that the author(s) has
not considered and that has been shown to work in resource allocations to
municipalities in Japan under similar circumstances. A recent paper by Amy
Catalinac, Bruce Bueno de Mesquita and Alastair Smith tests a formal model in
which variance in centrally-provided local municipality resources is a function
of a tournament among municipalities within a larger legislative district.
Perhaps future research might compare the impact of the tournament structure
with the local city council effects in Brazil as studied here."

\vspace{.2cm}

\noindent \textbf{Response:}

\begin{quote}

We agree with Reviewer 2 that testing the proposed mechanism is an exciting
project for a follow-up article. In the Discussion section (page 29), we write:
"\textit{Second, future research could evaluate whether additional legislators
help mayors gain access to federal or state transfers in countries where
resources are centrally provided, such as Japan. As suggested by
\citet{catalinac2020tournament}, in these situations the national government can
create a tournament between municipalities and provide resources to those which
offer more electoral support. Scholars might want to analyze whether this
tournament structure interacts with the mechanisms we suggest in this paper}."

\vspace{.2cm}

We would like to thank Reviewer 2 for their helpful comments and suggestions.

\end{quote}

# Reviewer 3 Comments and Responses

\noindent 1) Reviewer 3 writes: "The central story rests on some assumptions
that I would like to see justified. Do mayor's generally want to spend more,
and the city council holds them back? Why? I wonder about this assumption. Why
does increasing the size of the city council usually result in pro-mayor
additions? Under what conditions might increasing the size of the council bring
in more diverse viewpoints or more polarized policy stances? What role is
partisanship playing? It seems like it is mainly a proxy for agreement between
councilors and the mayor, but do different parties have strongly different
policy preferences? Does that matter for the results presented here?"

\vspace{.2cm}

\noindent \textbf{Response:}

\begin{quote}

We collapse two concerns from Reviewer 3 since they both speak to core topics in
our paper. In Table 6 of the Supplementary Materials (page 21), we show that an
additional legislator usually does not lead to significant increases in public
expenditure. Therefore, our results speak more about the allocation of resources
than about government spending. In our section titled "Background on the
Brazilian Case" (page 14), we clarify why increasing the city council size often
results in a pro-mayor legislator. We write (page 15): "\textit{Brazil also
adopts a unique formula to allocate seats, which combines the Hare quota and the
D’Hondt method and strongly favors larger parties \citep{nicolau2015sistemas}.
This poses a major obstacle for small parties to reach the electoral quorum, so
they often join party coalitions to improve their chances of winning legislative
seats \citep[68]{Ames2001}. In Brazil, mayors are usually members of the largest
coalitions \citep{poulsen2018political}, thus there is an incentive for
candidates to side with the mayor before and after the election}." We mention
that information again on page 22, footnote 12: "\textit{The effect of city
council size on the number of mayor-aligned politicians stems from the rounding
rule adopted by the proportional representation system in Brazil. The rounding
rule favors the coalitions with more votes. Most of the time, the coalition of
the elected mayor is the one that receives more votes.}".

We did not find evidence that electing an additional legislator improves
descriptive representation in the city council. For instance, we see no
significant increase in the number of female or non-white legislators, as shown
in Table 4 (page 27). Finally, we show in the table below that the mayors'
parties are not driving our results. This result is also available on our GitHub
repository. We ommitted it from the Supplementary Materials to conserve space.

\begin{table}[!htbp] \centering 
\caption{Councilors from Major Parties} 
\label{appcouncilparty} 
\resizebox{\textwidth}{!}{
  \begin{tabular}{@{\extracolsep{\fill}}l c c c c} 
  \\[-1.8ex]\hline \\[-1.8ex] 
  & \multicolumn{1}{c}{\textbf{Councilors from PT}} & \multicolumn{1}{c}{\textbf{Councilors from PSDB}} & \multicolumn{1}{c}{\textbf{Councilors from PMDB}} & \multicolumn{1}{c}{\textbf{Councilors from PFL (DEM)}} \\ 
  \hline \\[-1.8ex] 
  LATE & -0.01 & -0.05 & -0.01 & -0.01 \\ 
  & (0.02) & (0.05) & (0.03) & (0.02) \\ 
  N Left & 5179 & 5179 & 5179 & 5179 \\ 
  N Right & 343 & 343 & 343 & 343 \\ 
  Eff N Left & 375 & 152 & 267 & 231 \\ 
  Eff N Right & 212 & 121 & 172 & 159 \\ 
  BW Loc Poly & 13.99 & 6.8 & 10.58 & 9.63 \\ 
  BW Bias & 22.64 & 11.373 & 16.803 & 15.647 \\ 
  \hline \\[-1.8ex] 
  \multicolumn{5}{l}{$^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1} \\ 
  \end{tabular} }
\end{table}

\end{quote}

\noindent 2) Reviewer 3 writes: "Finally, I wonder about the scope conditions.
The evidence for the patterns presented from Brazil are very convincing, but
I'd like to see more discussion of why these patterns hold in Brazil and how
they might be different/ similar elsewhere."

\vspace{.2cm}

\noindent \textbf{Response:}

\begin{quote}

We entirely agree with Reviewer 3 that we need to discuss the scope conditions
of our findings in more detail. We acknowledge that the external validity of
our results might be conditional on the contextual characteristics of our
study. We have rewritten two paragraphs in the Discussion (page 29) to highligh
the implication of our results. We write: "\textit{Our findings
have implications that extend well beyond Brazil. A general form of our
argument is that it is easy for a politician to negotiate with members of her
in-group \citep{abbink2019, g9030057}. Specifically, several
countries have separate executive and legislative powers, such as the federal
level of presidential systems, state governments, and local city councils. As
these institutions are influenced by political allegiances, the dynamics we
expose here may be generalizable. Depending on legislative support, changing
the legislature size could increase the chance of electing politicians that are
not aligned with the executive, hence decreasing welfare. Despite the
particular effects on political support and welfare, the logic described in
this paper would remain relevant for empirically accessing the effect of
enlarged legislatures on welfare. Our findings may also help scholars
understand why other countries in Latin America and Asia can generate effective
governance despite having strong executives and large coalitions
\citep{pereira2012surprising}.}".

In the next paragraph (page 29), we indicate some of the limitations of our
findings and encourage scholars to conduct further research on the topic. We
write: "\textit{This paper opens several questions for future analysis. [...]
Third, scholars might want to investigate whether our findings remain valid
under different conditions. It is possible that the mechanisms we present here
are non-linear, with additional legislators having a positive or negative
impact on welfare depending on certain electoral characteristics. Finally, it
remains unclear how other legislative features affect citizen well-being. For
example, city councils may vary in size, monetary compensation for councilors,
committee structure, and internal power structure. Understanding how
legislature size interacts with other features would improve our knowledge
about how local and national legislatures impact public services.
These questions are crucial for institution design and welfare in developing
democracies.}"

We would like to thank Reviewer 3 for their comments and suggestions.
\end{quote}

\newpage