### 2022-05-28 article/article.Rmd
reduce word count article and memo

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: 
date: \today
abstract: "How does legislature size impact public service provision? Despite the importance of institutional design for democratic governance, the effect of legislative features on citizen welfare remains little understood. In this paper, we use a formal model to show that increasing legislature size improves public goods delivery. We argue that changes in bargaining costs depend on whether additional legislators share the executive's party affiliation: More opposition members reduce the equilibrium public goods provision, while more government-aligned members increase it. We test this theory by exploiting sharp discontinuities in city council size in Brazil. We show that an additional city councilor has a 91% chance of belonging to the mayoral coalition, and this significantly improves primary school enrollment and infant mortality rates. To explore possible mechanisms, we surveyed 174 former city councilors and analyzed 346,553 bills proposed between 2005 and 2008. This paper has implications for the design of representative institutions."
abstractspacing: double
keywords: legislature size; legislative bargaining; local public goods provision; regression discontinuity design; Brazil
jelcodes: H41; H80; I31; I38
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
# Knitr options
knitr::opts_chunk$set(fig.pos = "H") # holds figure position
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

# Introduction
\label{sec:intro}

Legislative institutions are crucial for citizen welfare and public service
provision in democratic countries. Historically, the separation of powers and
the birth of representative governments placed legislatures at the center of
the political stage. Today, these institutions check and veto other political
powers, determine taxation, and oversee public programs at all levels of
government \citep{auriol2012optimal, poulsen2018political, stasavage2011states,
Weingastetal1981}.

Regardless of eventual differences in electoral rules, social composition, or
internal procedures, all legislatures share one common feature: They are
collective bodies. As such, the number of representatives has important
consequences for the whole polity. Legislature size has a significant effect on
the bargaining power between state branches, public expenditures, and
representation \citep{crain1979cost, freire2021legislature,
gerring2020population, PrimoSnyder2008, rogers2002free, Weingastetal1981}.
However, how legislature size impacts social welfare remains less understood.
Larger legislatures may reduce political particularism and increase overall
government spending, which leads to more public goods
\citep{gerring2020population, GilliganMatsusaka2001, Weingastetal1981}. But
adding representatives to a legislature may exacerbate collective action
problems, thus hindering service provision \citep{gottlieb2019countervailing,
tsebelis2002veto}. Therefore, what is the effect of larger legislatures on
public service delivery?

In this paper, we propose a theory based on executive leadership and party
politics to explain how city council size impacts local public goods provision.
We incorporate partisanship and mayoral first-move advantage to a
\citet{baron1989bargaining} bargaining model. Mayors make the initial policy
proposal, which consists of investments in public goods provision, rents for
personal gains, and a vector of side-transfers for councilors. If mayors fail
to secure majoritarian support from councilors, a reversal stage occurs, in
which the decision leaves the mayor's office and councilors propose the public
investments. In this bargaining process, costs come from the side-transfers
required to secure a majority among councilors. We show that in a model without
partisan concerns, bargaining costs always increase with larger city councils,
which in turn lower public goods provision and rents. However, partisanship
decreases bargaining costs when the chances of electing government-aligned
councilors are sufficiently high. Lower bargaining costs release funds for
mayors to invest in public goods. In this case, a larger city council may add
more members to the government coalition, thus reducing the mayor's bargaining
costs and increasing the equilibrium public goods provision.

We test these theoretical predictions by exploiting exogenous variation in city
council size in Brazil. Brazilian cities could freely determine their
legislature size, but in March 2004 the Supreme Electoral Court unexpectedly
reinterpreted the 1988 Constitution and added one councilor to each set of
47,619 inhabitants. Around these cutoffs, the decision represented an exogenous
change in city council size, allowing us to investigate the effects of
increasing the number of local legislators in Brazil.

We find that larger city councils changed the composition of mayoral coalitions
and the number of appointed bureaucrats in the municipality. The extra city
councilor had a 91 percent chance of belonging to the mayor's pre-electoral
coalition. With a larger coalition, mayors face lower bargaining costs and are
able to boost public investments with fewer constraints. Each additional
legislator also leads to 104 politically appointed bureaucrats, and research
shows that political appointees improve public service provision and enhance
accountability in Brazil \citep{toral2019benefits}.

To provide further evidence for our mechanisms, we surveyed 174 former
councilors who served during the 2005-2008 term, the period we analyze in this
paper. They confirm that mayors indeed use bureaucratic appointments and
political favors to secure legislative support. As our formal model shows,
bargaining with councilors requires offering transfers to the representatives.
Councilors also demand several job appointments, which are regarded as an
effective means to boost their reelection prospects and improve municipal
welfare. These findings are in line with quantitative results from a novel
dataset we built with 346,553 laws passed in 63 municipalities within ten
thousand inhabitants away from the cutoffs. While all municipalities mention
public goods in their legislation, those with larger city councils have 15
percent more proposals about public service delivery.

Finally, we show that the increase in legislature size significantly impacts
municipal education and health care. Increasing the city council size by one
legislator lowers infant mortality by 2.01 children per one thousand infants
born and reduces postneonatal mortality by 0.90 children per one thousand
infants who survived their first 28 days. Larger councils also increase
enrollment by 2.58 children in elementary education classrooms without reducing
school quality. These services are crucial for citizen welfare and result from
eased bargaining between the mayor and the city council.

This paper contributes to four strands of the literature. First, this paper
dialogues with a large literature on multilateral bargaining. Repeated models
of multilateral bargaining started with \citet{baron1989bargaining}, and
scholars have recently analysed how endogenous recognition power, the size of
the required majority, and partisanship impact these results
\citep{agranov2020persistence, baranski2020determinants,
calvert2005legislative, choate2019partisan}. Here we include the mayor's first
move and political partisanship to Baron and Ferejohn's
(\citeyear{baron1989bargaining}) and investigate the often-overlooked effect of
legislature size on bargaining equilibrium strategies.

Second, this paper contributes to the literature on the effects of larger
legislatures on service provision. Following the seminal contribution by
\citet{Weingastetal1981}, the "law of $1/n$" posits that larger chambers
provide incentives for legislators to provide targeted public goods to local
supporters at the expense of the entire polity \citep{freire2021legislature}.
This question has become more relevant since legislature size has increased
consistently over time \citep{gerring2020population}. The corollary of the "law
of $1/n$" is that increasing legislature size promotes excessive public
spending, regardless of the partisan makeup of the legislature
\citep{GilliganMatsusaka2001}.\footnote{\citet{PrimoSnyder2008} show that
legislature size may result in positive or negative expenditures, depending on
factors such as the degree that local public goods can exclude non-locals or
deadweight costs associated with taxation.} However, the literature on the "law
of $1/n$" rarely mentions the relationship between public goods provision and
welfare. To our knowledge, this paper is one of the first to study how the
number of local legislators affects citizen well-being.

Third, our article dialogues with the growing literature on municipal partisan
politics. Most papers in this field estimate the effects of the mayor's party
on political outcomes. For instance, \citet{de2016mayoral} and
\citet{gerber2011mayors} find differences in governance when comparing
Democratic and Republican mayors. Specifically about Brazil,
\citet{gouvea2021partisanship} show that social spending increases with
left-wing mayors. Using discontinuities created by a constitutional amendment
that reverted the thresholds we study here, \citet{frey2021legislature} shows
that changes in council size lower the support for mayor-backed candidates. Our
paper reinforces the view that partisanship matters locally, even in countries
with weak ideological linkage.

Finally, this paper also contributes to the literature on the causes and effects of
legislature sizes. While most papers on legislature size analyse the trade-off
between representation and efficiency \citep{frederick2008people,
jacobs2015explaining, Stigler1976}, recent articles have
explored the effects of legislature size on lawmaking and partisanship in
further detail. \citet{kirkland2014chamber} shows that larger legislatures make
it harder for politicians to know and collaborate with each other, so they rely
more heavily on party cues and committee structures.
\citet{jacobs2015explaining} show that population shifts tend to increase the
number of local legislators, but legislature size tends to decrease during
economic crises. \citet{marland2019fewer} questions the latter mechanism,
showing that political elites shrink legislatures mostly to signal to the
public the priority of their austerity agenda. In this article, we show that
council size affects welfare by changing the partisan composition of the
legislatures.

# A Theory of Legislature Size, Legislative Bargaining, and Welfare
\label{sec:theory}

Institutional design has a lasting impact on social welfare
\citep{acemoglu2012nations, buenodemesquita2005logic}.
One issue that all polities face is how to determine the optimal size of their
legislative bodies. The literature on this topic is extensive, and modern
attempts to address the legislature size problem started in the early
eighteenth century with the Federalists versus Anti-Federalists debates
\citep{Stigler1976}.

The literature on legislature size highlights that the number of legislators
causes tension between representation and efficiency
\citep{frederick2008people, jacobs2015explaining, taagepera1972size}. On the
one hand, larger legislatures improve representation and facilitate the
electoral success of ethnic and political minorities
\citep{allenrepresentation, gerring2020population}. On the other hand, better
representation is often countered by the decline in efficiency in larger
legislatures \citep{Weingastetal1981}. In large councils, it is harder for
members to collaborate with each other without some type of partisanship or
formal legislative structure \citep{kirkland2014chamber}. Larger chambers also
have more potential veto players, who may decrease policymaking efficiency
\citep{tsebelis2002veto}. In this sense, balancing representation and
efficiency is key to find the optimal legislative composition
\citep{auriol2012optimal, taagepera1972size}.

However, two questions have often been overlooked in these debates. First, does
legislature size matter in contexts where legislative bodies are not pivotal
decision-makers? Second, what is the ultimate effect of legislature size on
welfare? These gaps are likely generated by the predominance of the "law of
1/n" scholarship in this literature, which focuses mainly on the relationship
between legislature size and overspending. Since \citet{Weingastetal1981}, 30
papers have tested the effects of legislature size on government expenditures
\citep{freire2021legislature}, and the only article which examined its effects
on welfare found negative results \citep{lewis2019legislature}.
\citet{frey2021legislature} analyzes the Brazilian case after the 2009 change
in legislature size and finds a null effect in service provision from 2012 to
2016. Yet, \citet{GilliganMatsusaka2001} find that in the US legislature size
is associated primarily with higher spending on geographically concentrated
benefits, such as education and highways. These results suggest that the
legislature size and welfare nexus needs a dedicated theoretical and empirical
assessment.

## Model Setup

In a mayor-council setting, consider a strategic interaction between a mayor
($M$) and $N \geq 3$ city councilors indexed in $i \in \{1, 2, \cdots, N \}$.
The mayor has the prerogative of proposing a vector of policies to be voted by
the city councilors. The provision of public goods is then a bargain between
the mayor and the city council. Policy proposals are a combination of public
goods provision $g$, rents for the mayor $r$, and a vector of transfer for city
councilors \textbf{x}.

The city council votes the mayor's proposal and the proposal is implemented if
most councilors accept it. Otherwise, a reversal policy takes place. Reversal
policy in this context means that the mayor's proposal failed to secure the
majority of council votes. The decision process then leaves the mayor's office
and is transferred to the council. The types of reversal policies are crucial
for our argument, as they change the relative strength of the council when
legislators bargain with the mayor. Here we investigate two types of reversal
policies: A baseline reversal with no parties and a hybrid mechanism with
partisan and non-partisan concerns. The differences between these two
mechanisms are that the non-partisan councilors only care about rents, while
partisan councilors care about both rents and policy.

The mayor must convince at least half of the city councilors to support her
policy proposals. Similar to \citet{buenodemesquita2005logic}, we assume that
the selectorate, that is, the individuals who have a say in the mayor's policy
proposals, are the $N$ city councilors. The winning coalition, which is the
minimum number of members that must be convinced to implement the policy, is
equal to half of the selectorate ($N/2$). City councilors are motivated by
policy concerns $p$ and transfers received from the mayor or the councilor
nominated as the proposer \textbf{x}. We leave the definition of these
transfers open, as this allows us to analyze how mayors combine transfers with
other incentives to pressure city councilors. Examples of transfers are
granting electoral favors, funding clientelistic brokers, spending in areas in
which the councilor has political or personal interests, or providing portfolio
(municipal offices) and public jobs within the bureaucracy. The use
of transfers to influence the councilors' choices generate a bargaining cost of
$C_G$. Bargaining costs vary according to the reversal mechanism and the size
of the legislature.

Finally, after learning the bargaining costs, the mayor provides a level of
public goods provision $g$ and pockets rents $r$. Public goods provision helps
the mayor to get reelected. Rents are for the mayor's direct consumption and do
not contribute to her electoral success. This makes the mayor's expected utility a
sum of the gains from rents and the benefits from reelection. Both the
utilities from rents $u(.)$ and the probability of reelection $\pi(.)$ are
concave functions, meaning that more rents or public goods increase the utility
at a decreasing rate. The probability of reelection is multiplied by the
benefit from holding office $B_M > 0$. This benefit captures the tangible and
intangible gains that the mayor receives from holding the public office. We
assume that the benefits are high enough to rule out an equilibrium with only
rents and transfers, which is equivalent to saying that mayors strongly prefer
to be reelected. If the mayor's policy proposal is approved, then the expected
utility for the mayor is:

\[
\mathbb{E}U(r, g) \ = \ u(r) + B_M\pi(g)
\]

The policy choices of the mayor are subject to municipal budget constraints.
The municipality has $R > 0$ resources and cannot run debts, meaning that the
budget must be balanced. We also assume that there are enough resources for the
mayor to govern.

Let $C_G$ be the bargaining costs. The budget balance constraint requires that
the offers the mayor makes to the city councilors must satisfy the following
inequality:

\[
r + g + C_G \ \leq \ R
\]

The expected utility for the city councilors depends on the type of reversal
policy. In the baseline non-partisan reversal, when the council rejects the
mayor's proposal, a reversal stage starts with resources diminishing by a
factor of $\delta \in (0,1)$ and the random selection of one councilor. The
selected councilor becomes the next proposer. If the new proposal is accepted,
it is implemented. If councilors reject the offer again, the budget shrinks by
$\delta$ and another city councilor is nominated to propose. The process
repeats until a proposal is finally accepted.

We now consider a situation that combines both partisan and non-partisan
legislature concerns. Each councilor has a party affiliation. Party
affiliations are mutually exclusive, and if a city councilor is aligned with
the mayor's party, then she belongs to the government coalition $G \subset \{1,
2, \cdots, N \}$. Otherwise, the city councilor belongs to the opposition $O =
\{1, 2, \cdots, N \} \setminus G$. The mayor's policy generates a political
value of $p > 0$ for the councilors, and whether it adds or subtracts from the
councilor's utility depends on the party affiliation of the councilor. If
councilors reject the mayor's proposal, no public goods are provided ($p=0$).
Moreover, in equilibrium, we assume that city councilors would prefer to
receive transfers instead of providing public goods.

The solution concept we use in this model is the subgame perfect Nash
equilibrium. This game requires that the strategies follow a Nash equilibrium
in each subgame. We find the solution using backward induction. Since the model
may have infinitely repeated proposals, any strategy can be shown to be a Nash
equilibrium. Thus, a natural assumption is to extend the equilibrium concept to
require the equilibrium to be stationary. A stationary subgame perfect
equilibrium requires that, at each given point of the game, if a politician
accepts an offer at time $k+1$, she should take the same offer at time $k$.
Stationarity gives us a method to find a proposal that would be accepted at any
stage of the game. Hence, we can characterize a sequence of offers for each
point $k$ in time. For the mayor, the optimal offer is at $k=0$, representing a
no delay in policy implementation. We assume that players have no dominated
strategies, even when there is insufficient support for the policy.

For the solution to be of general interest, we solve the bargaining costs for
the expected number of opposition and government-aligned city councilors. This
is because the solution for each fixed number of opposition and government
councilors does not help us understand the overall relationship between council
size and bargaining costs.

## The Mayor's Decision Stage

Solving the game by backward induction requires us to start with the
councilors' strategies. To conserve space, assume that we solved the
game for the councilors and found the equilibrium bargaining costs $C_G(N)$
when there are $N$ city councilors. We now derive the optimal rents ($r$) and
public goods provision ($g$) proposed by the mayor. The mayor benefits from
public goods provision, as it increases the chances of her reelection. However,
she prefers to invest as minimum as necessary in public goods and extract the
remaining resources as political rents. The mayor's objective is to maximize
her expected utility, subject to the municipal budget constraint.

\[
\begin{aligned}
\max_{r, g} \quad & u(r) + B_M\pi(g)\\
\textrm{s.t.} \quad & r + g + C_G(N) \leq R
\end{aligned}
\]

In equilibrium, the budget binds, and the first-order condition for optimal
public goods provision indicates that the marginal costs of providing public
goods is equal to the marginal benefits of reelection:

\[
u'(R - g - C_G(N)) = B_M\pi'(g)
\]

The optimal level of public goods provision equates the marginal gains accrued
from rents with the marginal benefits from reelection.

\begin{proposition}\label{prop1}
Public goods provision increases with legislature size if bargaining costs decrease with legislature size.
\end{proposition}

All proofs are in section A.1 of the Supplemental Materials. The proof's
intuition involves finding what happens to the mayor's marginal utility when we
increase both $g$ and $N$. As $N$ increases discretely on a complete lattice,
we use monotone comparative statics to derive the conditions for increasing
differences \citep{ashworth2006monotone}. Increasing differences mean that
higher $N$ leads to higher equilibrium $g$.

Proposition \ref{prop1} provides our first empirically testable hypothesis:
\textit{if bargaining costs decrease when legislature size increases,
then public goods provision increases when legislature size increases}. Now
we study each reversal mechanism to determine the condition under which the
bargaining costs decrease when legislature size increases.

## Baseline: Non-Partisan Legislative Bargaining

In the non-partisan reversal mechanism, we assume that when the council rejects
the mayor's office, a reversal stage starts with the random selection of one
councilor. This formulation is similar to the closed rules proposal in
\citet{baron1989bargaining}. To find a stationary subgame perfect Nash
equilibrium, suppose that there were $k-1$ rejections, and the game is at the
$k$-th stage. A councilor accepts the proposer's offer if, and only if, to
accept the offer at $k$ is better than or equal to wait until the next stage
$k+1$. If the offer is $x_i$, then:

\[
x_i \geq \dfrac{1}{N}\left[ \delta^{k+1}R - \dfrac{N}{2}x_i \right] + \left[ 1 - \dfrac{1}{N} \right]\left[\dfrac{1}{2} x_i \right]
\]

We place the offer on the left-hand side. There are two components on the
right-hand side. The first is the amount that the councilor $i$ gets when she
is the proposer. It is equal to the chance that she is recognized as the
proposer times the budget in the next round minus the offers she makes to
convince half of the councilors. The second part represents the gains if the
councilor rejects the current offer but still gets an offer in the following
round. It is equal to the chance that she is not recognized as the proposer
times the chance that she receives a transfer times the transfer amount. Note
that the offer the councilor makes as the proposer is the same as the offer she
wants to receive. This is because city councilors are exchangeable. The
solution is symmetric for all councilors receiving an offer (this means that,
without loss of generality, we could have dropped the $i$ in the solution).
After some algebra, the offer $x_i$ must be greater than or equal to:

\[
x_i \geq \dfrac{2 \delta^{k+1}R}{2N + 1} \equiv \underline{x}(k, N)
\]

The proposer always offers the minimum required to get the proposal approved.
In this case, the offer at any given stage $k$ is going to be equal to
$\underline{x}(k, N)$.

Proceeding backwards, at the mayor's proposal stage $k=0$ the mayor is going to
offer $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the
councilors. In this context, the bargaining costs for the mayor are:

\[
C_G(N) \ = \ \dfrac{N}{2}\left[\dfrac{2 \delta R }{2N + 1}\right] \ = \ \dfrac{\delta R N}{2N + 1}
\]

\begin{proposition}
In the baseline non-partisan reversal, bargaining costs always increase when legislature size increases.
\end{proposition}

The intuition for the results is that every additional councilor demands more
resources from the mayor. Thus, more councilors represent higher bargaining
costs. And if the bargaining costs increase with legislature size, then any
increase in public goods provision associated with legislature size has to come
from a different mechanism. Below we demonstrate that partisanship is the key
to understand how this relationship works.

## Partisan and Non-Partisan Bargaining

In a reversal mechanism that incorporates partisan and non-partisan
motivations, a city councilor aligned with the government favors the mayor's
offer if:

\[
x_i \geq \dfrac{2\delta R}{2N+1} - p
\]

An opposition politician, on the other hand, favors the mayor's offer if, and
only if:

\[
x_i \geq \dfrac{2\delta R}{2N+1} + p
\]

We decompose these costs into two components, which are the costs in terms of
rents and the costs (or benefits) from political alignment. In this context,
the bargaining costs depend on what the mayor offers for councilors in her
coalition versus what she provides to members of the opposition. As the ex-ante
chance of an opposition member being elected is equal to $1-\gamma$, after
taking the weighted averages these costs become:

\[
C_G(N) \ = \ \dfrac{N}{2}\left( \dfrac{2\delta R}{2N+1} - p \right) + \left(\dfrac{N}{2} - \gamma N \right)2p
\]

\begin{proposition}
In the reversal mechanism that incorporates partisanship, if $\gamma \geq \dfrac{1}{p}\left[ \dfrac{1}{(2N+1)(2N+3)} \right] \equiv \underline{\gamma}$, then bargaining costs decrease when legislature size increases.
\end{proposition}

The intuition for the proof is the following. As partisanship influences the
costs and benefits of legislators, more government-aligned councilors make it
easier for the mayor to govern. When the partisan benefit is higher than the
additional cost that a legislator generates, adding a councilor will benefit
mayors if the councilor has a higher chance of sharing the same party as the
mayor. Therefore, a higher probability of electing a legislator aligned with
the mayor reverts into lower bargaining costs.

Also, the threshold $\underline{\gamma}$ that captures the probability of
electing a government supporter decreases along with the political value of
policy $p$ and the size of the legislature $N$. This means that intensive
partisanship and larger legislatures make it easier to satisfy the electoral
threshold.

## Main Hypotheses

Our model provides two empirically testable hypotheses. Consider a municipality
with a city council comprised of $N$ legislators. Then the city increases the
size of the council to $N+1$. Thus, in a partisan setting:

\begin{enumerate}[labelindent=*, 
style=multiline, leftmargin=*, 
label=H\arabic*.]
  \item Bargaining costs decrease with legislature size when the chances of electing a government partisan are sufficiently high.
  \item Public goods provision increases if bargaining costs decrease with legislature size.
\end{enumerate}

# Background on the Brazilian Case
\label{sec:background}

In 2004, Brazil comprised 5,560 municipalities. According to the 1988 Brazilian
Constitution, each city must provide health care, primary education,
transportation, and infrastructure to its citizens. Municipalities have the
authority to enact local laws and collect taxes on housing and services.
However, their room for fiscal manipulation is limited. On average, municipal
taxes account for only 5\% of the local budgets, while mandatory transfers from
state and federal governments correspond to the remaining 95\%. This reduces
the ability of municipal legislators to effectively increase local spending,
which makes improvements in local services dependent only on the composition of
their local investments.

With regards to their political organization, municipalities have one mayor and
one city council. Citizens choose their representatives by direct vote in the
same election. Mayors oversee the provision of public services and propose laws
and amendments to the tax code. Proposals submitted by the mayor are subject to
the approval of the city council. Councilors, in turn, have two primary duties.
First, they should discuss and vote on legislation put forward by the mayors or
by their peers. Second, they oversee the public budget and sanction mayors who
do not comply with fiscal legislation. Councilors and mayors also provide an
array of public and personal services for their constituents, such as helping
voters access public hospitals, facilitating school enrollment, and even paying
their voters' utility bills \citep{lopez2017legisladores,
nichter2011electoral}.

Until 2004, there was no specific regulation on the size of city councils. The
Brazilian Constitution set only broad guidelines about the number of council
seats in each municipality, allowing cities to determine how many legislators
their local chambers should have. This led to severe imbalances in the
representation ratio across the country. For instance, the Nova Russas
municipality had 30,009 inhabitants and 21 councilors, or one legislator per
1,429 citizens. In contrast, the city of Sorocaba had only 15 councilors for
528,735 inhabitants, that is, one legislator per 35,249 citizens. In 2003, this
malapportionment problem gained national visibility when Mira Estrela, a small
municipality with only 2,651 residents, reduced its city council from 11 to 9
seats because of fiscal considerations. The change motivated a legal dispute
that reached the Brazilian Supreme Court. The Court favored the city's decision
and ruled that the size of every local council should follow population thresholds.

Following this ruling of the Supreme Court, in March 2004 the Electoral Court
established a series of population cutoffs for all municipal legislatures. The
Court specified that cities with up to one million residents should have
between 9 and 21 legislators. The number of seats was allocated as follows.
Each city starts with nine councilors, then adds one legislator for every
47,619 inhabitants until their councils reach 21 members. This threshold
implies that cities with a population between 571,428 and one million should
all have 21 legislators. Conversely, cities between one and four million
inhabitants should have at least 33 councilors, and this number increases up to
41 members following the previous formula of one additional legislator per
47,619 inhabitants.

Brazil also adopts a unique formula to allocate seats, which combines the Hare
quota and the D'Hondt method and strongly favors larger parties
\citep{nicolau2015sistemas}. This poses a major obstacle for small parties to
reach the electoral quorum, so they often join electoral coalitions to improve
their chances of winning legislative seats \citep[p. 68]{Ames2001}. In Brazil,
mayors are usually members of the largest coalitions
\citep{poulsen2018political}, thus there is an incentive for candidates to side
with the mayor before and after the election.

These factors make Brazil the ideal testing ground for our theory. Although an
endogenous decision motivated the Electoral Court ruling, the population
cutoffs created sharp discontinuities in the distribution of local council
seats. Cities close to the cutoffs had very similar characteristics, but those
just above the threshold gained a new legislator. Moreover, cities could not
self-select into any group, as their population estimates were calculated using
the 2003 projections by the Brazilian Census Bureau (IBGE). This change in
council size also happened only seven months before the election, making it
impossible for candidates to adjust their service provision strategies for the
2004 election. Therefore, the ruling allows us to study the effect of larger
legislatures on public service provision during the 2005-2008 term while
holding all else constant.

# Variables and Data Sources
\label{sec:sources}

We use four groups of city-level variables in our models. The first group
includes pre-treatment control variables that were not affected by the 2004
council size resolution. We add them to the estimates to improve the balance
between treated and non-treated units. The covariates are: 1) the number of
council seats in the previous term; 2) city population; 3) city GDP; 4) the
proportion of low-income families in each municipality. The data come from the
Supreme Electoral Court and the 2000 Brazilian census.

Second, we look into municipal data, city councilors' characteristics, approved
legislation, and an online survey with 174 former city councilors that served
during 2005--2008 term to investigate potencial political mechanisms. The
municipal data contain: 1) the number of councilors who belong to the mayor's
pre-electoral coalition; 2) the number of appointed bureaucrats; 3) the
councilor's gender and race; 4) the competitiveness of each city council seat;
5) the proportion of laws proposed by the city council that was approved. The
variables come from the Supreme Electoral Court (TSE), the Senate Legislative
Data Service (InterLegis), and the Brazilian Bureau of Statistics (IBGE).

Third, we collected 346,553 bills from cities within ten thousand inhabitants
from the 2004 population thresholds. We selected 63 out of 202 municipalities
whose voting records were available online. We evaluated proposal types,
discriminating between local public goods, oversight, and other legislative
activity. Examples of the local public goods are investing in a public health
clinic, requesting school staff to accommodate an extra child, renovating
sewage, and fixing potholes on a given street.\footnote{Our theory does not
differentiate between clientelism and broad service provision, as it is hard to
disentangle between the two in local legislation. A voter may request a
particular service, such as fixing a pothole. Attending this request can be
interpreted as a clientelistic practice. Still, the provision of the service
generates non-excludable benefits to the adjacent neighborhood.} Oversight
bills are accountability and transparency mechanisms meant to keep public
services provision in check. Other legislation comprises bills that do not fit
in the previous categories, such as changes in street names, motions to honor
citizens or groups, and legislative and internal city council procedures. We
provide further information about the legislation dataset and the city council
survey in sections A.13 and A.14 of the Supplemental Materials.

The last set of variables quantifies social welfare. Our focus is on the impact
of legislature size on education and health care, which are the two most
important public services the municipal government should provide to citizens.
According to the Brazilian Constitution, municipalities must spend 25\% of their
budget on education and 15\% on health care. Consequently, the Ministry of
Health and the Ministry of Education provide readily available, fine-grained
data on health care and education at the municipal level, which we leverage in
this study. We measure access to education with the average enrollment in
primary (K--4) municipal public schools. We assess education quality with the
Education Development Index (Índice de Desenvolvimento da Educação
Básica--IDEB), which the Ministry of Education has issued biannually since
2007. The index is a weighted average of student scores in Portuguese and
Mathematics (0--10), multiplied by the harmonic mean of public school
promotion rates in a given year (0--100). To test the effect of the 2004
ruling on health care, we collect data on infant mortality per thousand
children born and on postneonatal infant mortality, which is the death of
infants who survived more than 28 days, but died before reaching one year
of age. The source of all health care variables is the Brazilian Ministry
of Health Data Center (DataSUS).

# Empirical Strategy
\label{sec:empstrat}

We employ a regression discontinuity design (RDD) using the population
thresholds established by the 2004 Supreme Electoral Court resolution as
treatment indicators. Figure \ref{fig:runningvarplot} displays the distribution
of municipalities within each city council size. These population thresholds
remained in effect until 2009 when Congress amended the Brazilian Constitution.
This study covers 2005 to 2008, which constitutes the full mayoral and
city-council term following the 2004 Brazilian elections.

Our identification strategy relies on three assumptions. First, no municipality
should be able to self-select into each side of the discontinuity. Second, as
we estimate a sharp RDD, the council size should increase precisely as the law
mandates. Third, this design assumes that the pre-treatment variables,
collected before the 2003 Supreme Federal Court decision, were not affected by
the 2004 ruling on council size. Moreover, for all the actual models estimated
here, we fitted placebo cutoffs to check whether the estimation technique leads
to consistent results. The placebo cutoffs are the dashed lines in Figure
\ref{fig:runningvarplot}.

\vspace{.2cm}

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=6.8in,angle=0]{../figures/distrSeats.pdf}}
      \caption{Distribution of Municipalities by Population and City Council Size}
      \bigskip
      \label{fig:runningvarplot}
    \end{minipage}
  \end{center}
\end{figure}

In section A.3 of the Supplemental Materials we assess distribution imbalances
by running the \citet{McCrary2008} and \citet{cattaneo2019simple} tests. The tests
show no distribution imbalances.

For the second assumption, running RDD in multiple thresholds may lead to
inconsistent estimates when there are differences in the distribution of cases
around the cutoffs \citep{cattaneo2016interpreting}. In the dataset, there are
twelve discontinuities (from 9 to 21 legislators), but if we pool all the
discontinuities together with no correction, we find an increase in council
size of 1.63 councilors (see the first estimate in Table \ref{validitycheck}).
This imbalance occurs because when we combine all the discontinuities, the
model implies that municipalities below the 47,619 cutoffs (9 to 10 councilors)
are comparable to changes in towns right above the 571,428 cutoffs (20 to 21
council members).

In section A.5 of the Supplemental Materials we show by simulation that adding
controls, especially the variables responsible for the multiple threshold
assignments (in our case, the population in 2003), improves the consistency and
efficiency of the estimates. Relying on this fact, we add five controls to the
estimates: Population in 2003; GDP per capita; the number of seats in 2000;
year; and a dummy for Northeast municipalities. The reason for the inclusion of
first variable is the multiple threshold assignments. We add GDP per capita
because wealthier cities tend to have better public services. We include the
number of seats in 2000 because some municipalities could have experienced a
change provoked by the previous council size, thus confounding our estimates.
The Northeast dummy was added to control for the fact that the Lula
administration invested heavily in the region from 2003 until 2010, improving
several development indicators \citep{de2015politica}. We included an indicator
for year to improve efficiency and control for seasonal effects. Moreover,
adding controls improves efficiency in RDD \citep{calonico2019regression} and
also enhances the consistency in the multiple thresholds RDD.

For the last assumption, there should be no changes in pre-treatment
covariates. These pre-treatment covariates are measured before the primary
outcomes. As the threshold rule selected by the Brazilian Electoral Court was
unpredictable, we should not detect any variation before the thresholds were in
place.

\singlespacing

\begin{table}[H] \centering 
  \caption{Research Design Validity Check} 
  \label{validitycheck} 
    \scalebox{0.9}{
  \begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{4}{c}{\textbf{Panel A: Validity Check --- Treatment Effect on City Council Size}} \\ \cline{2-5}
   & \multicolumn{1}{c}{\small{(1)}} & 
 \multicolumn{1}{c}{\small{(2)}} & 
 \multicolumn{1}{c}{\small{(3)}} & 
 \multicolumn{1}{c}{\small{(4)}} \\

 & \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} & 
 \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} \\ 
  & \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} & 
  \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} \\
  & \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
\\[-1.8ex] 
 LATE & 1.63$^{***}$ & -0.31$^{**}$ & 1.00$^{***}$ & -0.00 \\ 
  & (0.51) & (0.14) & (0.0004) & (0.0000) \\ 
\hdashline \\[-1.8ex]   
 N Left & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} \\ 
Eff N Left & \multicolumn{1}{c}{199} & \multicolumn{1}{c}{638} & \multicolumn{1}{c}{49} & \multicolumn{1}{c}{886} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{385} & \multicolumn{1}{c}{51} & \multicolumn{1}{c}{477} \\ 
BW Loc Poly & 8.717 & 6.700 & 3.008 & 8.629 \\ 
BW Bias & 13.620 & 12.114 & 5.008 & 13.466 \\ 
\midrule 
 & \multicolumn{4}{c}{\textbf{Panel B: Validity Check --- Pre-Treatment Variables}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(5)}} & 
 \multicolumn{1}{c}{\small{(6)}} & 
 \multicolumn{1}{c}{\small{(7)}} & 
 \multicolumn{1}{c}{\small{(8)}} \\
  & \multicolumn{1}{c}{\small{Number of}} & 
 \multicolumn{1}{c}{\small{Population}} & 
 \multicolumn{1}{c}{\small{GDP}} & 
 \multicolumn{1}{c}{\small{\% of Poverty}} \\ 
  & \multicolumn{1}{c}{\small{Seats 2000}} & 
  \multicolumn{1}{c}{\small{2000 Census}} & 
  \multicolumn{1}{c}{\small{2000 Census}} &
   \multicolumn{1}{c}{\small{2000 Census}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
 LATE & 0.17 & -2057.60 & 0.06 & -9.72 \\ 
  & (0.67) & (1926.33) & (0.18) & (6.27) \\ 
  \hdashline \\[-1.8ex]
 N Left & \multicolumn{1}{c}{5178} & \multicolumn{1}{c}{5131} & \multicolumn{1}{c}{5131} & \multicolumn{1}{c}{5131} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} \\ 
Eff N Left & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{219} & \multicolumn{1}{c}{200} & \multicolumn{1}{c}{194} \\ 
Eff N Right & \multicolumn{1}{c}{147} & \multicolumn{1}{c}{157} & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{142} \\ 
BW Loc Poly & 8.970 & 9.368 & 8.739 & 8.530 \\ 
BW Bias & 13.860 & 15.029 & 14.752 & 13.035 \\ 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $ ^{***} $ p $ < $ .01; $ ^{**} $ p $ < $ .05; $ ^{*} $ p $ < $ .1. RD local linear estimates using \citet{calonico2019regression} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population, GDP per capita, number of seats in 2000, year, and dummy for Northeast region. \textit{N Left} and \textit{N Right} represent the total number of observation in the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

\doublespacing

\newpage

Table \ref{validitycheck} displays the results of the validity checks. We
present the first-stage regressions at the top of the Table, with and without
covariates, for both real and placebo cutoffs. At the bottom, we estimate the
pre-treatment covariate balance. In sections A.6 to A.12 of the Supplemental
Materials, we run placebo checks for all the models using placebo cutoffs
within the real ones.

# Empirical Results
\label{sec:results}

To test the predictions from our formal model, we present four sets of
empirical results. First, we confirm the bargaining costs hypothesis and find
that the additional legislator also promotes more appointments of bureaucrats.
This aligns with the councilors' descriptions of the strategies mayors employ
to consolidate support from city councilors. Second, we show that
municipalities with an extra city councilor approve more legislation about
public goods provision in general and about education and health care in
particular. Third, we show that all these pieces of evidence result in
improvements in infant mortality and elementary school enrollment, two vital
services for welfare. Finally, we also find that our results are robust to
alternative channels for the influence of legislature size on welfare.

## Council Size, Bargaining Costs, and Job Appointments
\label{sub:bargcosts}

The main premise of our model is that larger legislature size produces lower
bargaining costs when the chances of electing government-aligned politicians
are sufficiently high. Moreover, the main bargaining chip between the executive
and the legislative is job appointments, even more so when the city councilors
are aligned with the mayor. A higher number of political appointees is
associated with better service provision in Brazil, as their tenure in office
depends directly on the survival of the politician supporting them
\citep{toral2019benefits}. Career bureaucrats pass through a rigorous
examination and cannot be fired once hired, which considerably constrains
politicians' pressure on them. Finally, city councilors who want to influence
policy may employ more staff members and assistants to their offices. This
could influence policymaking through city councilors putting pressure on the
bureaucracy, defying our hypothesized mechanism. Our theoretical expectations
for these regressions are that larger council size affects the political
alignment of city councilors and the number of politically appointed
bureaucrats. It should not affect the number of career bureaucrats or city
council assistants, as it is more valuable to have a political appointee that
politicians can pressure, working directly with voters. Table \ref{tab:bargain}
tests the first hypothesis about the nexus between city council size and
bargaining costs and explores the consequences of larger legislatures on the
municipal bureaucracy.

\vspace{.4cm}

\begin{table}[!htpb] \centering 
  \caption{Political Effects of Increasing the Number of Legislators} 
  \label{tab:bargain} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} &
 \multicolumn{1}{c}{\small{(3)}} & 
 \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Mayoral}} & 
 \multicolumn{1}{c}{\small{Num. Politically}} & 
 \multicolumn{1}{c}{\small{Num. Career}} & 
 \multicolumn{1}{c}{\small{Num. Councilor}} \\
 & \multicolumn{1}{c}{\small{Coalition Size}} & 
 \multicolumn{1}{c}{\small{Appointed Bureaucrats}} & 
 \multicolumn{1}{c}{\small{Bureaucrats}} & 
 \multicolumn{1}{c}{\small{Assistants}} \\[1ex]  \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
 LATE & 0.91$^{*}$ & 104.49$^{*}$ & 71.57 & 2.09 \\ 
  & (0.50) & (61.53) & (218.12) & (4.37) \\ 
    \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5179} & \multicolumn{1}{c}{15536} & \multicolumn{1}{c}{15531} & \multicolumn{1}{c}{5179} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{1027} & \multicolumn{1}{c}{1027} & \multicolumn{1}{c}{344} \\ 
Eff N Left & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{372} & \multicolumn{1}{c}{513} & \multicolumn{1}{c}{99} \\ 
Eff N Right & \multicolumn{1}{c}{162} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{388} & \multicolumn{1}{c}{101} \\ 
BW Loc Poly & 10.025 & 6.242 & 7.695 & 5.134 \\ 
BW Bias & 15.889 & 10.243 & 11.780 & 8.902 \\ 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

Column (1) shows a 91\% chance of the extra legislator belonging to the mayoral
coalition. The model hypothesis stated that the probability of electing a city
councilor who is aligned with the mayor is sufficiently high. Sufficiently high
can be ambiguous, but in any case, 91\% chance is a considerably large
quantity. This strengthens the political position of the mayor and diminishes
bargaining costs.\footnote{The effect of city council size on the number of
mayor-aligned politicians stems from the rounding rule adopted by the
proportional representation system in Brazil. The rounding rule favors the
coalitions with more votes. Most of the time, the coalition of the elected
mayor is the one that receives more votes.} Column (2) demonstrates that adding
one councilor increases appointed employees by around 104 extra employees per
additional city councilor. As expected, legislature size has a null effect on
the number of career bureaucrats (Column 3) and councilors' staff members
(Column 4).

## Approved Legislation and City Councilor's Survey
\label{sub:survey}

We built two novel datasets to provide evidence for our mechanism. First, we
ran a survey with former city councilors, asking how mayors consolidate their
coalition support within the city council. Second, we collected 346,553 laws
approved in 63 municipalities within ten thousand inhabitants away from the
city council size thresholds. We hand-coded 3,466 laws into four categories:
Public goods, oversight, education and health care-related legislation, and
other legislative duties. After hand-coding these laws, we used a Supporting
Vector Machine (SVM) algorithm to classify the remainder laws. From all of the
algorithms we tested, SVM achieved the highest performance.

\vspace{.2cm}

\begin{figure}[H]
  \begin{center}
  \caption{Mayoral Instruments for Securing Councilors' Support}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=5in,angle=0]{../figures/strategySupportMayors.pdf}}
      \bigskip
      \label{survelec}
    \end{minipage}
  \end{center}
\end{figure}

We assume that city councilors want to extract rents and be reelected. The
rents come from continuing in the repeated game, and the partisanship discounts
that benefit. Figure \ref{survelec} shows that 65.9 percent of councilors
mentioned that mayors use job appointments to consolidate their coalition
support. Moreover, 59.0 percent of the surveyed councilors affirmed that mayors
use personal favors, such as shifting resources to places where the councilors
have electoral support, to solidify their governing coalition. This provides
further evidence that mayors do offer city councilors the benefits they demand.
In this regard, granting favors works as rent transfers for councilors. Job
appointments are also convenient for the mayors because patronage in Brazil has
positive results in terms of welfare \citep{toral2019benefits}.

\begin{figure}[H]
  \begin{center}
  \caption{Legislation Approval Patterns in the Municipalities Closer to the City Council Size Thresholds}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=5in,angle=0,height=6in]{../figures/lawsApproved.pdf}}
      \bigskip
      \label{LEGAPPR}
    \end{minipage}
  \end{center}
\end{figure}

In Figure \ref{LEGAPPR}, we show that one additional city councilor increased
the per capita amount of laws on public goods by 15.0 percent. Also, the number
of laws that focus on health care and education increased by 19.7 percent.
Looking specifically into health care and education legislation, the proportion
of laws discussing public goods increased by 15.7 percent, showing that policy
proposals by both the mayor and the councilors shift toward local and citywide
public goods.

## The Effect of Legislature Size on Welfare
\label{sub:effects}

We now study the impact of legislature size on education and health care. Table
\ref{tab:outcomes} shows that increasing the city council by one legislator
generates welfare improvement in municipalities. Regarding health care,
increasing the legislature size decreases infant mortality in 2.01 deaths per
1,000 children born. This impact is significant in absolute terms and
represents an effect of 0.14 in standard deviations. Moreover, postneonatal
mortality decreases by 0.90 casualties per 1,000 births, which remains around
the 0.14 standard deviations improvement as infant mortality. Infant mortality
is a primary outcome for welfare, and it correlates with several other
indicators of well-being \citep{world2005health}. The fact that infant
mortality improved in the Brazilian case shows that the increased mayor's
support mechanism results in large welfare effects. As expected, placebo
regressions were statistically insignificant.

\begin{table}[!htpb] \centering 
  \caption{Legislature Size and Welfare} 
  \label{tab:outcomes} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Infant}} & 
 \multicolumn{1}{c}{\small{Postneonatal}} &
  \multicolumn{1}{c}{\small{School}} & 
  \multicolumn{1}{c}{\small{Education}} \\ 
   & \multicolumn{1}{c}{\small{Mortality}} & 
 \multicolumn{1}{c}{\small{Mortality}} &
  \multicolumn{1}{c}{\small{Enrollment}} & 
  \multicolumn{1}{c}{\small{Quality}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & -2.01$^{**}$ & -0.90$^{*}$ & 2.52$^{***}$ & -0.03 \\ 
  & (0.79) & (0.48) & (0.80) & (0.13) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{12299} & \multicolumn{1}{c}{5441} & \multicolumn{1}{c}{10156} & \multicolumn{1}{c}{8597} \\ 
N Right & \multicolumn{1}{c}{1029} & \multicolumn{1}{c}{672} & \multicolumn{1}{c}{686} & \multicolumn{1}{c}{669} \\ 
Eff N Left & \multicolumn{1}{c}{501} & \multicolumn{1}{c}{324} & \multicolumn{1}{c}{206} & \multicolumn{1}{c}{409} \\ 
Eff N Right & \multicolumn{1}{c}{381} & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{291} \\ 
BW Loc Poly & 7.527 & 7.564 & 5.425 & 9.05 \\ 
BW Bias & 11.95 & 11.01 & 10.58 & 15.107 \\ [1.0ex] 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

In terms of education, with the addition of a city councilor, enrollment in
elementary schools increased by 2.58 children on average per school. This
represents a change of around 0.20 standard deviations. Besides the influence
of the mayoral support, enrollment can also directly impact the councilor's
representation: Councilors can pressure the bureaucracy to accommodate a few
extra students. With support from the mayor, they can help voters get access to
schools, improving the welfare of voters. Education quality remains unchanged,
but with more students in the classroom, it could potentially decrease. Again,
the placebo regressions are insignificant.

## Robustness to Alternative Mechanisms

Other pathways may also explain the nexus between city council size and
welfare. First, improvements in representation could change the provision of
health care and education. Increasing legislature size without changing the
population makes councilors represent fewer people. This facilitates the
representation of minorities, as they have to gather fewer votes to elect a
city councilor. A minority representative can advocate for policies that would
improve the welfare of marginalized groups \citep{ChattopadhyayDuflu2004,
chin2011redistributive}. Diversity could also improve governance in
municipalities \citep{rugh2011provision}. For instance, more women in public
offices may shift service provision toward welfare expenditure
\citep{brollo2016happens, duflo2012women, funk2019representative,
hernandez2018women}. Therefore, an alternative hypothesis to explain the
increased service provision is that larger legislatures improve female and
minority representation in Brazil.\footnote{For a description of the barriers
faced by non-white politicians in Brazil, see \citet{janusz2021race}.}

Second, more seats could increase the election competitiveness, forcing
politicians to provide more services to remain competitive
\citep{besley2010political}. In a political system with around thirty parties
and a sample average of almost ten parties in the municipalities we study,
candidates per seat can increase exponentially. According to the Brazilian
electoral law, each party can sponsor 1.5 candidates per seat. If a party forms
a pre-electoral coalition, this number increases to two times the number of
seats \citep{coneglian2008lei}. Therefore, an alternative explanation for
better service provision is that a larger council size increases the
competitiveness of the election.\footnote{\citet{boulding2014political} and
\citet{gottlieb2019countervailing} show that more competition can increase
budgetary pressure and lower social policy, leading to bargaining inefficiencies.}

Finally, both the mayor and the legislature carry out the provision of
services. As in firms, an extra employee can increase the firm productivity
frontier by improving cooperation and synergistic outputs
\citep{holmstrom1982moral}. A similar argument would posit that the increase in
council size increases the capability for municipalities to provide public
services. As legislative productivity manifests itself as legislation, an
alternative hypothesis to explain the increased service provision is that more
legislation is being approved by the city council.

\begin{table}[!htbp] \centering 
  \caption{Robustness: Effects of Legislature Size Changes on Representation, Competitiveness, and Legislative Productivity} 
  \label{tab:robust} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Num. Female}} & 
 \multicolumn{1}{c}{\small{Num. Non-white}} &
  \multicolumn{1}{c}{\small{Candidates}} & 
  \multicolumn{1}{c}{\small{Prop. Laws}} \\ 
   & \multicolumn{1}{c}{\small{councillors}} & 
 \multicolumn{1}{c}{\small{councillors}} &
  \multicolumn{1}{c}{\small{Per Seat}} & 
  \multicolumn{1}{c}{\small{Approved Council}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & 0.22 & 0.60 & -0.19 & -0.02 \\ 
  & (0.38) & (0.85) & (1.11) & (0.08) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5183} & \multicolumn{1}{c}{239} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{3424} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{158} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{269} \\ 
Eff N Left & \multicolumn{1}{c}{195} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{179} & \multicolumn{1}{c}{227} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{132} & \multicolumn{1}{c}{141} \\ 
BW Loc Poly & 8.623 & 2.958 & 8.144 & 10.976 \\ 
BW Bias & 13.565 & 4.922 & 12.018 & 16.708 \\[1.0ex] 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

Table \ref{tab:robust} tests whether our results could stem from these
different mechanisms. Columns (1) and (2) test the hypotheses that better
representation of under-served groups could improve welfare. As we see, larger
city councils do not increase the representation of females or non-whites.
Column (3) shows that the competition per seat in municipalities with larger
city councils remain statistically equal to zero. Finally, Column (4) shows
that the proportion of legislation passed by the city council remains unchanged
when council size increases.

# Discussion

This paper shows that partisanship considerably decreases bargaining costs and
improves citizen welfare. The main providers of public goods are the mayors,
but they cannot do it alone. Legislatures provide a solid check to the mayor's
agenda. Bargaining with legislators in a country plagued by corruption can be
costly \citep{bertholini2017price}. Therefore, an extra ally in the city
council may ease the bargaining process and increase social welfare.

Increasing the council size results in gains in elementary school enrollment
and in lower infant mortality. Education quality remains unchanged, but
increasing enrollment without affecting quality should be interpreted as a
positive outcome. The analysis of the approved legislation shows that the
frequency of laws discussing local public goods increases in municipalities
above the council size thresholds. Finally, our survey data demonstrate that
councilors and mayors agree about appointing more civil servants to the public
administration.

\citet{Ames2001} argues that the Brazilian democracy is prone to deadlocks
because the executive has its hands tied by legislators who demand consistent
provision of private goods. Indeed, the most successful presidents after 1995
had to build extensive coalitions to secure legislative support
\citep{limongi1998bases, raile2011executive}. Despite evidence that side
payments cemented these coalitions, many social and economic policies have
since been implemented. For instance, education and health care notably
improved during the Cardoso and Lula administrations
\citep{acemoglu2012nations}. In contrast, presidents who challenged the
coalition system did not succeed. For instance, former President Rousseff was
impeached in her second term. Bolsonaro can be considered one of the most
ineffective presidents in Brazilian history.

Municipalities work similarly. In this case, the stakes are lower and the
negotiations involve only one legislative branch. But in a system such as
Brazil's, with more than thirty-five registered parties
\citep{samuels2014power}, mayors face substantial challenges to approve their
political agendas. Additional support in the council reduce governing costs and
improve public goods provision.

However, we do not argue that these benefits result from the absence of
corruption or clientelism. Both the formal model and empirical evidence
provided by \citet{britto2020corruption} show that corruption increases with
legislature size. As the model shows, city councilors may receive a side
transfer from the mayor to support her agenda. Mayors also divide the resources
saved in the bargain into public goods and rents. The rate for the division
depends on the marginal changes of the rents and public goods, but these two
components receive the extra resources. In some cases, the mayor's rents may be
primarily corruption, and the councilor's transfers can be clientelistic and
aimed at improving his or her reelection prospects.

Our findings have implications that extend well beyond Brazil. A general form
of our argument is that it is easier for a politician to negotiate with
in-group members \citep{abbink2019, alt2018giving}. Specifically, several
countries have separate executive and legislative powers, such as the federal
level of presidential systems, state governments, and local city councils. As
these institutions are influenced by political allegiances, the dynamics we
expose here may be generalizable. Depending on legislative support, changing
the legislature size could increase the chance of electing politicians that are
not aligned with the executive, hence decreasing welfare. Despite the
particular effects on political support and welfare, the logic described in
this paper would remain relevant for empirically accessing the effect of
enlarged legislatures on welfare. Our findings may also help scholars
understand why other countries in Latin America and Asia can generate effective
governance despite having strong executives and large coalitions
\citep{pereira2012surprising}.

This paper opens several questions for future analysis. First, if the mechanism
we suggest is indeed at work, one may ask what should be the city council's
checking prerogatives, as they significantly reduce the provision of public
services. The evidence so far is inconclusive, as \citet{poulsen2018political}
show that checks for the mayors may also improve service provision. Second,
future research could evaluate whether additional legislators help mayors gain
access to federal or state transfers in countries where resources are centrally
provided, such as Japan. As suggested by \citet{catalinac2020tournament}, in
these situations the national government can create a tournament between
municipalities and provide resources to those which offer more electoral
support. Scholars might want to analyze whether this tournament structure
interacts with the mechanisms we suggest in this paper. Third, scholars might
want to investigate whether our findings remain valid under different
conditions. It is possible that the mechanisms we present here are non-linear,
with additional legislators having a positive or negative impact on welfare
depending on certain electoral characteristics. Finally, it remains unclear how
other legislative features affect citizen well-being. For example, city
councils may vary in size, monetary compensation for councilors, committee
structure, and internal power structure. Understanding how legislature size
interacts with other features would improve our knowledge about how local and
national legislatures impact public services. These questions are crucial for
institution design and welfare in developing democracies.

\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}